### 2013-04-17 ordinal/response.Rnw
Final to ordinal ms prior to resubmission o Added YB97 statistic to response o Used apa6 class instead of apa class for the ms,   allowing for easy masking of authors
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@200 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[a4paper]{article}
\usepackage{a4wide}
\usepackage[english]{babel} % To avoid tildes
\SweaveOpts{engine = R, eps = FALSE, echo = FALSE, results = hide}

\begin{document}

\setkeys{Gin}{width=\textwidth}

\section*{Reviewer 1}

\begin{enumerate}
\item \emph{A comparison of the current approach with scenarios in which the
factor analyst uses Modification Indices (MIs) to free parameters. In practice
if one has, say, 5 age groups and a given parameter is found to lead to
significant increases in model fit when restricted across groups, it is quite
common to use MIs to detect the source(s) of the misfit and then to end up with,
say one parameter for age groups 1-3  and another parameter for age groups 4 \& 5
in a so-called partially invariant model (see Byrne et al., 1989). It would be
interesting to see how that approach fares in comparison to the approaches
described in the manuscript. I am also interested in reading the authors' ideas
about the use MI in ``ad hoc'' model fitting.}

MIs are quite related to the proposed tests: the MI described by, e.g., Sorbom
1989 is equivalent to a Lagrange multiplier test.  The maximum of Lagrange
multiplier tests (e.g., computed across all potential thresholds of V), on the
other hand, is within the family of statistics described by Merkle \& Zeileis
(2013).  However, as the reviewer states, MIs are typically applied only after a
significant likelihood ratio test has been obtained.  Our proposed statistics
can be used across multiple parameters, potentially in place of the likelihood
ratio test.  If the test is significant, one could use MIs or parameter
fluctuation plots such as those in manuscript Figures 3 \& 4 to determine the
model misfit.  We have added some brief detail on this relationship to the
``Tests for Continuous V'' section.

\item \emph{Given the widespread literature on ordinal factor analyses, it would
be worthwhile to link the current modelling approach to ordinal factor analysis.
Another major reason is that ordinal factor analysis naturally extends towards
more standard IRT models. Therefore, by bridging the gap from the current
modelling approach to ordinal factor analysis, the authors may also take steps
to integrate it with the literature in measurement invariance in the IRT
framework.}

We have added detail in the General Discussion on the extension of the proposed
tests to other factor analysis models and to IRT.  In particular, our proposed
tests can potentially extend to any psychometric model that assumes independent
observations and that uses ML estimation.  Additionally, we cite some work on
the application of similar tests to IRT.

\item \emph{Study of power under other violations of MI related to factor
loadings or measurement intercepts.}

We now acknowledge that the obtained simulation results may not generalize to
other measurement invariance violations or to other factor analysis models.  We
feel that this suggestion would require multiple simulations that would result
in a much longer paper, so we have not added new simulations to the manuscript.

\item \emph{Consider issues related to model comparison by means of AICs or BICs
and the like. This is particularly relevant in light of the fact that monotonic
violations of MI should be detectable in a more parsimonious manner than how it
is often done in practice with likelihood ratio tests with many DFs.}

If we understand this comment correctly, the reviewer is suggesting that we
compare the proposed statistics to the use of AIC or BIC (in addition to the
LRT).  Under this framework, one would, e.g., reject the hypothesis of
measurement invariance if the AIC for the ``full'' model were less than the AIC
for the restricted model.  We have added the AIC strategy to our simulation,
along with some other adjusted versions of the LRT (see Response 2.2).  The AIC
turns out to have particularly low power.
\end{enumerate}

\section*{Reviewer 2}

\begin{enumerate} \item \emph{On page 5 is the scaled sum of scores which is
normalized by the square root of n and the covariance matrix of the scores I.
This is fine theoretically, but what choice do you make in practice? In fact I
think that here the problem is model misspecification, other than the biasing
factor V. If the model used in ML is incorrect with respect to the number of
factors for instance, then this will result in a inconsistent estimate of I. Why
not use the sandwich by the late Halbert White (JASA, 1982 for instance)? This
is easy to calculate and will improve the limiting distribution of the test
statistics.  The choice of I is also missing in the simulations. Please be more
specific what you did there.}

We have added an extended discussion of choices for the covariance matrix of
scores, with references.  We have also clarified that, in the simulations, we
used the information matrix.  Because our simulation models were correctly
specified (except for the measurement invariance violation), we do not expect
the choice of I to play a large role there.

\item \emph{The simulations show great improvement over the standard LR test on
measurement invariance. And on page 10 the authors mention the results by
Bentler and Bonett (1980) on LR tests picking up deviations too quickly. I think
the alternative tests, adjustments of LR tests, like the one proposed in Yuan
and Bentler (JASA, 1997). There the LR is downweighted at rate $O_p(1/n)$, which
brings the results of the LR closer to the results of the authors' maxLM test. I
would like to see a comparison with this or maybe some other adjusted LR test.}

<<fig1, echo=FALSE, results=hide, fig=TRUE, height=10, width=10>>=
load("sim1.rda")
library("lattice")

lev0 <- c(   "LM_o","LRT","LRT_sb","AIC","YB97")
lev  <- c("maxLM_o","LRT","LRT_sb","AIC","YB97")
cl <- c(1, 4, 2, 3, 5)
pc <- c(1, 2, 3, 4, 5)
lt <- c(1, 1, 2, 2, 2)

sim1.tmp <- sim1[sim1$test %in% lev0,]
sim1.tmp$test <- factor(sim1.tmp$test, levels = lev0, labels = lev)
trellis.par.set(theme = canonical.theme(color = FALSE))
print(xyplot(power ~ diff | nlevels + nobs, group = ~ test, data = sim1.tmp,
  type = "b", xlab="Violation Magnitude", ylab="Power",
  key=list(text = list(lab = lev), lines = list(col = cl, lty = lt), points = list(col = cl, pch = pc)),
  col = cl, pch = pc, lty = lt))
@

<<fig2, echo=FALSE, results=hide, fig=TRUE, height=10, width=10>>=
load("sim2.rda")
sim2.tmp <- sim2[sim2$test %in% lev0,]
sim2.tmp$test <- factor(sim2.tmp$test, levels = lev0, labels = lev)
trellis.par.set(theme = canonical.theme(color = FALSE))
print(xyplot(power ~ diff | nlevels + nobs, group = ~ test, data = sim2.tmp,
  type = "b", xlab="Violation Magnitude", ylab="Power",
  key=list(text = list(lab = lev), lines = list(col = cl, lty = lt), points = list(col = cl, pch = pc)),
  col = cl, pch = pc, lty = lt))
@

We implemented three adjusted LRTs in our simulations: (i) the Yuan-Bentler 1997
adjustment suggested by the reviewer, (ii) the use of AIC, and (iii) the
Satorra-Bentler adjustment with correction for difference tests.  While these
adjustments downweight the LRT associated with individual models (e.g.,
comparing an individual model to a saturated model), the impact of the
adjustments on difference tests is less clear.  Further, the Yuan-Bentler
adjustment is seemingly designed for situations where the model is estimated via
WLS, and there are no available corrections for difference tests.  We suspect
that these issues are problematic, as our simulation results indicated that
these statistics are worse than the usual LRT in nearly all conditions.
We have included the resulting graphs for Simulation~1 and Simulation~2 above
(but not in the manuscript).
\begin{itemize}
  \item Both graphs are organized in the same manner as the graphs in the paper,
    except that they contain different statistics: the ordinal maxLM test
    ($\mathrm{maxLM}_o$), the Likelihood ratio test (LRT), the
    Satorra-Bentler LRT ($\mathrm{LRT}_\mathrm{sb}$), the Yuan-Bentler
    statistic (YB97), 
    and the use of AIC (AIC). The first two are presented in the
    manuscript while the last three are only shown here.
  \item It can be seen that $\mathrm{LRT}_\mathrm{sb}$ is virtually identical to
    the plain LRT in situations with small number of groups $m$ and/or large
    sample sizes $n$ (in particular throughout Simulation~2) while for the large $m$/small
    $n$ scenarios, the problems of the LRT are even amplified.
  \item The YB97 statistic is problematic at smaller $n$ and at large
    $m$, generally resulting in 
    very high Type I error rates.  We suspect that this is because the
    statistic is based on models fits via WLS, which is generally
    known to require very large $n$ to be reliable.  The extra
    parameters from large $m$ would then seem to require even larger
    $n$.  However, 
    even in the Simulation~2 situations where the Type I error rate is
    true, the performance of the statistic is similar to that of the plain LRT.
  \item Finally, the AIC has relatively low power throughout Simulation~1. This is
    because using AIC corresponds to using the LRT with an increased critical value
    here (because the models being compared are nested with relatively large differences in degrees
    of freedom).
\end{itemize}
Hence, we feel that the additional discussion of these modifications do not add
substantial new insights and so have omitted them from the paper (and only
mention them very briefly). But we are willing to reconsider if the AE or
reviewers find the results worthwhile.
\end{enumerate}

\section*{Reviewer 3}

\begin{enumerate} \item \emph{The section entitled `Theoretical Detail' (page 4)
starts with the background for the score-based test statistics. However, from
lines 18 onwards, there is a section on measurement invariance (starting with
`In studying measurement invariance, we are essentially testing...' up to line
19 on page 5, just before the subsection `Tests for Continuous V'). I would
suggest to put the section on measurement invariance testing in a separate
(sub)section, and present this first, before providing the background for the
test statistics.}

We have moved around these sections, so that measurement invariance comes before
the theoretical discussion.

\item \emph{For readers unfamiliar with case-based scores, it may be useful to
have a table with artificial values for some $B(\theta)_{ij}$ values; say for 20
subjects, and 4 parameters; in an additional column, there could be ordered
continuous values for a V variable; in the next column, the V values could be
binned to get an ordinal variable; referring to this table, it would be trivial
to illustrate how the values of eg DM, CvM, maxLM, etc are computed...}

We agree that this would be useful for some readers, but we think that it would
take up more space than it is worth.  For example, the calculation of ordinal
test statistics would require both calculations within each bin and unique
scaling (by $(i/n) \cdot (1 - i/n)$) for each bin.  While these calculations are not
difficult, they also consume space to write them out (especially if we included
the non-ordinal statistics).  Additionally, besides the calculation of test
statistics, it might be equally useful to show how the scores $s(\theta; x_i)$ can
be used to calculate $B(\theta)$.  This implies that we would need both a table of
$s(\theta; x_i)$ and a table of cumulative scores $B(\theta)$.  We feel that these
ideas are represented more concisely and intuitively via their equations.

\item \emph{Related to 2), I would suggest adding (perhaps in an appendix) some
R code illustrating how (for example, for the values in the aforementioned
table) the p-values can be computed for the various test statistics}

We have added some practical discussion of p-value computation to the ``Choice of
Test'' subsection of the General Discussion.  We have also posted all the R code
for replicating our results (p-values and otherwise) on our R-Forge page (listed
in the ``Computational Details'' section).  We do not include code within the
manuscript because the p-values associated with some of the statistics are
computed via simulation (i.e., there is not a closed-form solution for critical
values associated with the test statistic), and much of the underlying R code is
already contained within R's strucchange package.
\end{enumerate}

All minor points have been addressed, with thanks to the reviewer.

\end{document}