### 2013-07-08 frontiers/testprops1.Rnw
add writing files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@212 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Editing Rules:                                                            %
%1. Ed's comment and questions:\readme{}                                   %     
%2. Ting's comment and questions:\seeme{}                                  % 
%3. When comments and questions are resovled, the corresponding \readme or %  
%   \seeme will be removed by Ting.                                        %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[doc]{apa}
\usepackage{graphicx,epsfig,amsmath,alltt,setspace,bm}
\usepackage[english]{babel}

\SweaveOpts{engine = R, eps = FALSE, echo = FALSE, results = hide}

\title{Tests of measurement invariance via stochastic processes:
  {S}ome properties}
\twoauthors{Ting Wang and Edgar C.\ Merkle}{Achim Zeileis}
\twoaffiliations{University of Missouri}{Universit\"{a}t Innsbruck}

\abstract{
We study a family of 
recently-proposed measurement invariance tests that are derived from
stochastic processes.  As applied to factor analysis and other
psychometric models, the tests are advantageous in that (i) they do
not require advance definition of subgroups of cases that violate
measurement invariance; (ii) they can potentially identify specific
model parameters that violate measurement invariance; and (iii) they
do not require a model of additional complexity to be estimated.  
Because the tests have been applied to only a small number of
psychometric examples, we conduct a series of simulation studies that provide
a detailed examination of test properties.  These studies allow us
to study (details here).  The studies indicate that.}

\acknowledgements{This work was supported by National Science
  Foundation grant SES-1061334. 
  Correspondence to. 
  Email: }
\shorttitle{Measurement invariance test properties}
\rightheader{Measurement invariance test properties}

\newcommand{\argmax}{\operatorname{argmax}\displaylimits}
%% for internal use
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}
\newcommand{\seeme}[1]{\emph{\marginpar{SEEME} (#1)}}

\spacing{1}

\begin{document}
\maketitle

<<preliminaries1>>=
## packages
library("lavaan")
library("strucchange")
library("mvtnorm")
library("lattice")

## auxiliary code
source("simall.R")
source("mz.R")

## convenience function for plotting boundaries without color
get_boundary <- function(obj, fun, trim = 0) {
  bound <- fun$boundary(0:obj$nobs/obj$nobs)
  bound <- fun$computeCritval(0.05, ncol(obj$process)) * bound
  bound <- zoo(bound, time(obj$process))
  if(trim > 0) bound <- head(tail(bound, - floor(trim * obj$nobs)), - floor(trim * obj$nobs))
  return(bound)
}

## convenience function for applying tests to real data
info_full <- function(x, ...) solve(vcov(x) * nobs(x))
@

Parameter invariance is a fundamental assumption for measurement model in 
developing psychometric scales. Violation of it could happen in any estimated parameter.  
In order to fix the psychometric scale, we need to specify the violating parameter.  
Moreover, in practice model misspecification is an inevitable issue to a certain extent 
and may interfere with parameter invariance specificity. A family of tests derived from 
stochastic process has been recently proposed to identify specific parameters impacted 
by measurement invariance violation. However, these tests have only been used to a limited 
type of parameter violation. It is this paper's intent to provide a detailed examination 
of the tests' performance under practical research scenarios.

In the following section, we first briefly review the 
theoretical framework of proposed statistical tests. Subsequently, we study the tests'
ability through simulations that mimic practical research. 
Finally, some suggestion on the tests' use in practice is provided.

\section{Simulation 1}

In Simulation 1, we examined the extent to which the proposed tests
are sensitive to the specific model parameter that violates measurement
invariance.  If, say, a factor loading violates measurement
invariance, it is plausible that this violation impacts other
parameters such as factor covariances or other loadings 
associated with the factor in question.  Thus, the goal of the
simulation was to examine the extent to which the proposed tests
attribute the measurement invariance violation to the parameters that
are truly in violation.

\subsection{Methods}
To examine these specificity issues, we generated data from a
two-factor model with three indicators each (the same model used in
Merkle and Zeileis, \citeyearNP{MerZei13}).  The measurement invariance
violation occurred in one of three parameters: the factor loading
associated with the first indicator $\lambda_{11}$, the unique variance associated
with the first indicator $\epsilon_{11}$, or the factor covariance $\sigma_{12}$.  
We then tested for measurement invariance in seven subsets of parameters: the three
individual parameters noted above, all six factor loadings $\lambda_{11}--\lambda_{62}$, 
all six unique variances $\epsilon_{11}--\epsilon_{66}$. The path diagram is shown in  
Figure~\ref{fig:path}.

\begin{figure}
\caption{Path diagram representing the factor analysis model 
         used for simulations.}
\label{fig:path}
\includegraphics{PathDiagram}
\end{figure}

Power and Type I error were examined across three sample sizes
($n=100,200,500$) and 17 magnitudes of invariance violations. These
violations were based on the auxiliary variable $V$: individuals below
the 50th percentile of $V$ deviated from individuals above the 50th
percentile  by $d$ times the parameters' asymptotic standard errors
(scaled by $\sqrt n$), with $d=0,0,25,0.5,...,4$. 
For each combination of sample size ($n$) $\times$ violation
magnitude ($d$) $\times$ violating parameter, 5000 datasets were
generated and tested.

\subsection{Results}
Full simulation results are presented in Figure~\ref{fig:sim11res} to
Figure~\ref{fig:sim13res}; all three test statistics exhibited very
similar results, so the figures include only results for the CvM statistic.
Figure~\ref{fig:sim11res} displays power curves as a function of 
violation magnitude for the first
factor loading $\lambda_{11}$, with panels reflecting the parameters 
being tested and lines reflecting $n$.
Figures~\ref{fig:sim12res} and Figure~\ref{fig:sim13res} display 
similar power curves when the factor covariance 
$\phi_{12}$ and error variance $\epsilon_{11}$ violate measurement
invariance, respectively.

From these figures, one generally observe that the tests isolate 
the parameter violating measurementinvariance.  Additionally, the tests 
have somewhat-higher power to detect measurement invariance violations 
in the factor loading and factor covariance parameters, as opposed to the error variance
parameter.  Finally, simultaneous tests of all factor loadings or all error
parameters result in decreased power, as compared to the situation
where one tests only the violating parameter.  This occurs because, in
testing a subset of parameters (only one of which violates measurement
invariance), we are effectively dampening the signal of a measurement
invariance violation.

In summary, we found that the proposed tests can attribute
measurement invariance violations to the correct parameter.  This
provides evidence that, in practice, one can have confidence in the
tests' abilities to locate the measurement invariance violation.  Of
course, this statement is qualified by the fact that, in this
simulation, the model was correctly specified.  In the following
simulations, we examine the tests' performance when the model is
misspecified.

%% Example of how to call a simulation in R:
<<sim1>>=
## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1090)

if(file.exists("sim1.rda")) {
  load("sim1.rda")
} else {
  ## Arguments lead to small-scale simulation:
  sim1 <- simulation(nobs=c(100,200,500), diff=seq(0,4,.25), parms=c("loading","var","error"))
  save(sim1, file="sim1.rda")
}

## Add to column names so that plots look nicer
sim1$nobs <- paste("n=", sim1$nobs, sep="")

#Not sure how to express all loadings and all error terms in greek letters.
greekExprList<-expression(lambda[11],epsilon[11],sigma[12],lambda[11]--lambda[62],
                          epsilon[11]--epsilon[66])
greekpar<-as.expression(greekExprList)

@

%% Example of how to create a figure of results
\begin{figure}
\caption{Simulated power curves for Simulation 1.Violation Parameter is loading}
\label{fig:sim11res}
\setkeys{Gin}{width=\textwidth}
<<sim11res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey1 <- simpleKey(unique(sim1$nobs), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff | pars, group = ~ nobs, data = sim1,
             subset = diff <= 4 & parms == "loading" & test=="CvM" &  pars %in% 
             c("1","7","13","1:6","7:12"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,2),type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey1,
             as.table=TRUE))
@
\end{figure}

%% Example of how to create a figure of results
\begin{figure}
\caption{Simulated power curves for Simulation 1.Violation Parameter is covariance}
\label{fig:sim12res}
\setkeys{Gin}{width=\textwidth}
<<sim12res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
print(xyplot(power ~ diff | pars, group = ~ nobs, data = sim1,
             subset = diff <= 4 & parms == "var" & test == "CvM" &  pars %in% 
             c("1","7","13","1:6","7:12"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,2),
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey1,
             as.table=TRUE))
@
\end{figure}

\begin{figure}
\caption{Simulated power curves for Simulation 1.Violation Parameter is error terms}
\label{fig:sim13res}
\setkeys{Gin}{width=\textwidth}
<<sim13res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
print(xyplot(power ~ diff | pars, group = ~ nobs, data = sim1,
             subset = diff <= 4 & parms == "error" & test == "CvM" &  pars %in%
             c("1","7","13","1:6","7:12"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,2),
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey1,
             as.table=TRUE))
@
\end{figure}

%###########################################################################################
\section{Simulation 2}
In Simulation 2, we examine the extent to which the results of
Simulation 1 are robust to model misspecification. Specifically, we
generate data from the factor analysis model used in the previous
section, but with adding an extra loading from the second factor to the first
indicator (of the first factor).  The estimated model lacks this extra
loading, however. The goal of this simulation is to examine the proposed statitics'
power and spicificity under model misspecification.

\subsection{Method}
Measurement invariance violations could occur in
each of the three parameters from Simulation 1 (factor loading,factor covariance,unique
variance), and they could also occur in the extra,unmodeled loading. 
Sample size and magnitude of measurementinvariance violation were manipulated 
in the same way as they were in Simulation 1.The tested parameters are the same as Simulation
1,too.

\subsection{Results}
Results are presented by the paramter that violated measurement invariance.
Figure~\ref{fig:sim21res} displays results when the unmodeled loading violates 
measurement invariance. One can generally observe that 
the first loading and error variance parameter showed high measurement invariance 
violation simultaneously, while factor covariance did not demonstrate violation.

Figure~\ref{fig:sim22res} displays results when the loading $\lambda_{11}$ violates
measurement invariance. Even though lacking one loading associated with the first indicator, 
CvM statistic showed enough power to detect the first loading violation specifically. 
Similar pattern of specificity and power were observed when the violation parameter 
was $\sigma_{12}$ (Figure~\ref{fig:sim23res}) and $\epsilon_{11}$ (Figure~\ref{fig:sim24res}).

In summary, the proposed test statistics appear robust to unmodeled loading parameters,so long
as the umodeled parameter does not violate measurment invariance. If the unmodeled parameter 
dose violate measurement invariance, however, then the tests assign this violation to other 
parameters that do not violate measurement invariance. The impacted parameters include the error
variance and other loadings associated with the manifest variables that had the unmodeled 
loading. Thus, prior to testing measurement invariance, it is important to study the extent to
which the hypothesized model includes all parameters of importance.

%% Example of how to call a simulation in R:
<<sim2>>=
## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1090)

if(file.exists("sim2.rda")) {
  load("sim2.rda")
} else {
  ## Arguments lead to small-scale simulation:
  sim2 <- simulation(nobs=c(100,200,500), diff=seq(0,4,.25), parms=c("extra",
                    "extra+loading", "extra+var","extra+error"))
  save(sim2, file="sim2.rda")
}

## Add to column names so that plots look nicer
sim2$nobs <- paste("n=", sim2$nobs, sep="")

@

%% Example of how to create a figure of results
\begin{figure}
\caption{Simulated power curves for Simulation 2.Violation Parameter is extra loading}
\label{fig:sim21res}
\setkeys{Gin}{width=\textwidth}
<<sim21res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey2 <- simpleKey(unique(sim2$nobs), points = TRUE, lines = TRUE)
print(xyplot(power ~ diff |pars , group = ~ nobs, data = sim2,
             subset = diff <= 4 & parms == "extra" & test == "CvM" & pars %in%
              c("1","7","13","1:6","7:12"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,2),
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey2))
@
\end{figure}

%% Example of how to create a figure of results
\begin{figure}
\caption{Simulated power curves for Simulation 2.Violation Parameter is loading, 
         but model misspecified}
\label{fig:sim22res}
\setkeys{Gin}{width=\textwidth}
<<sim22res, fig=TRUE, height=7, width=7>>=
print(xyplot(power ~ diff | pars, group = ~ nobs , data = sim2,
             subset = diff <= 4 & parms == "extra+loading" & test == "CvM" & pars %in% 
             c("1","7","13","1:6","7:12"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,2),
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey2))
@
\end{figure}

%% Example of how to create a figure of results
\begin{figure}
\caption{Simulated power curves for Simulation 2.Violation Parameter is var, 
         but model misspecified}
\label{fig:sim23res}
\setkeys{Gin}{width=\textwidth}
<<sim23res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
print(xyplot(power ~ diff | pars, group = ~ nobs , data = sim2,
             subset = diff <=4 & parms == "extra+var" & test == "CvM" & pars %in%
             c("1","7","13","1:6","7:12"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,2),
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey2))
@
\end{figure}

%% Example of how to create a figure of results
\begin{figure}
\caption{Simulated power curves for Simulation 2.Violation Parameter is var, 
         but model misspecified}
\label{fig:sim24res}
\setkeys{Gin}{width=\textwidth}
<<sim24res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
## 
print(xyplot(power ~ diff | pars, group = ~ nobs , data = sim2,
             subset = diff <=4 & parms == "extra+error" & test == "CvM" & pars %in% 
             c("1","7","13","1:6","7:12"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,2),
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey2))
@
\end{figure}

%###########################################################################################

\section{Simulation 3}
In this section, we are aiming to test the extent to which the results of Simulation 1 
are robust to percentage of violation group impact. Specifically, we generate data from 
the factor analysis model used in Simulation 1 but manipulating the percentage of violating 
individuals. The goal of this simulation is to examine the proposed three statistics'
power and spicificity in different percentage violating individuals.

\subsection{Method}
The percentage of violation group is manipulated from 10\% to 90\% by 10\% in 
each simulation. Violation magnitude is fixed to 3. Sample size is fixed to 500. 
The violation parameters are the same as Simulation 1, including factor loading, 
factor correlation and error term. 5000 datasets were generated and tested.

\subsection{Results}
Simulation results were shown in Figure~\ref{fig:sim31res} 
to Figure~\ref{fig:sim33res}. The highest power exhibited 
when the percentage is 0.5 for all three violation paramter, as shown in Simualition 1 when 
group divided by half. As the percentage decrease or increase from 0.5, the power decreases.
However, the decreasing pattern is not symmetric. Specifically, when the violation percentage
is approaching to 0.1, the power decreases more rappidly and down to lower than 0.1 when the 
violation percentage is 10\%. Meanwhile, the power decrease less dramatically as the violation
percentage approaching to 0.9. As for the difference among these three statistics, we noticed 
that supLM which takes the maximum over individuals is the most robust to violation percentage.
The power remains to 0.2 as violation percentage reaches 0.9, while dmax and CvM drop to near 
0.1. This observation is consistent with the statistics' property. dmax or CvM takes the 
maximum or sums over the parameters and individuals, respectively. While supLM sums over 
parameters then take the maximum over individuals. Therefore, when the violation percentage is
small, this statistics is more resistent to the contamination produced by aggregation across 
individuals. 
However, when there are more than one parameter violates the measurement invariance, the 
difference pattern among these three statistics might change.We will explore this problem in 
simulation 4.

\seeme{The asymptotic pattern should be interpreted with the parameter value. I write some 
  in the discussion part. Basically, if most individuals' parameter value is larger, 
  the power will be larger.}
\seeme{I think the natural question would be when more than one paramter violate in different
  individual percentage, how the three statistics would perform. So these two simulations would
  be looked as exploration of the difference among these three statisitcs.}

%% Example of how to call a simulation in R:
<<sim3>>=
## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1090)

if(file.exists("sim3.rda")) {
  load("sim3.rda")
} else {
  ## Arguments lead to small-scale simulation:
  sim3 <- simulation(nobs=c(100,200,500), diff=seq(0,4,.25), parms=c("loading","var","error"))
  save(sim3, file="sim3.rda")
}

## Add to column names so that plots look nicer
sim3$test <- paste("", sim3$test, sep="")
levels(sim3$perc)<-c(".05",".1",".15",".2",".25",".3",".35",".4",".45",".5",".55",
                     ".6",".65",".7",".75",".8",".85",".9",".95")

@

%% Example of how to create a figure of results
\begin{figure}
\caption{Simulated power curves for Simulation 3.Violation Parameter is loading}
\label{fig:sim31res}
\setkeys{Gin}{width=\textwidth}
<<sim31res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey3 <- simpleKey(unique(sim3$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ perc | pars, group = ~ test, data = sim3,
             subset = diff ==3 & parms == "loading"  & nobs == 500 & pars %in%
             c("1","7","13","1:6","7:12") & perc %in% c(".1",".2",
               ".3",".4",".5",".6",".7",".8",".9"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,2),
             type = "b", xlab="Violation Percentage",ylab="Power", key=mykey3))
@
\end{figure}

%% Example of how to create a figure of results
\begin{figure}
\caption{Simulated power curves for Simulation 3.Violation Parameter is var}
\label{fig:sim32res}
\setkeys{Gin}{width=\textwidth}
<<sim32res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey3 <- simpleKey(unique(sim3$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ perc | pars, group = ~ test, data = sim3,
             subset = diff ==3 & parms == "var" & nobs == 500 & pars %in% 
             c("1","7","13","1:6","7:12") & perc %in% c(".1",".2",
               ".3",".4",".5",".6",".7",".8",".9"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,2),
             type = "b", xlab="Violation Percentage",ylab="Power", key=mykey3))
@
\end{figure}

%% Example of how to create a figure of results
\begin{figure}
\caption{Simulated power curves for Simulation 3.Violation Parameter is loading}
\label{fig:sim33res}
\setkeys{Gin}{width=\textwidth}
<<sim33res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey3 <- simpleKey(unique(sim3$test), points = TRUE, lines = TRUE)

print(xyplot(power ~ perc | pars , group = ~ test, data = sim3,
             subset = diff ==3 & parms == "error" & nobs == 500 & pars %in% 
             c("1","7","13","1:6","7:12") & perc %in% c(".1",".2",
               ".3",".4",".5",".6",".7",".8",".9"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,2),
             type = "b", xlab="Violation Percentage", ylab="Power", key=mykey3))
@
\end{figure}

%###########################################################################################

\section{Simulation 4}
In Simulation 4, we are aiming to test the extent to which the results of Simulation 1 
are robust to multiple group violation.Specifically, we generate data from the factor 
analysis model used in Simulation 1 but adding one more group which violates the measurement 
invariance by a different degree. The goal of this simulation is to examine the proposed 
statistics'power and spicificity in multiple group violation condition.

\subsection{Method}
The violations were still based on the auxiliary variable $V$, but there are two cutoff 
points for $V$. Therefore,individuals is divided into three groups.For simplicity reason,
individuals were assigned to these three groups equally. Group difference were $d/2$ times 
the parameters' asymptotic standard errors(scaled by $\sqrt n$), with $d=0,0,25,0.5,...,4$. 
For each combination of sample size ($n$) $\times$ violation magnitude ($d$) $\times$ violating 
parameter, 5000 datasets were generated and tested.

\subsection{Results}
Simulation results were displayed in Figure~\ref{fig:sim41res} to Figure~\ref{fig:sim43res}. 
The test specificity were retained as in two group setting.However, the powers are generally 
lower than the corresponding two group setting as shown in Simulation 1.

In summary, when there is more than one group violating measurement invariance, the proposed 
test statistic could remain specificity and enough power to test the particular violating 
parameter.

%% Example of how to call a simulation in R:
<<sim4>>=
## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1090)

if(file.exists("sim4.rda")) {
  load("sim4.rda")
} else {
  ## Arguments lead to small-scale simulation:
  sim4 <- simulation(nobs=c(100,200,500), diff=seq(0,4,.25), parms=c("loading","var","error"))
  save(sim4, file="sim4.rda")
}

## Add to column names so that plots look nicer
sim4$nobs <- paste("n=", sim4$nobs, sep="")
## Not fully happy with this, but haven't figured out how
## to get greek letters in xyplot panel labels.

@

%% Example of how to create a figure of results
\begin{figure}
\caption{Simulated power curves for Simulation 4.Violation Parameter is loading}
\label{fig:sim41res}
\setkeys{Gin}{width=\textwidth}
<<sim41res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey4 <- simpleKey(unique(sim4$nobs), points = TRUE, lines = TRUE)

print(xyplot(power ~ diff | pars, group = ~ nobs, data = sim4,
             subset = diff <= 4 & parms == "loading" & test=="CvM" & pars %in%
             c("1","7","13","1:6","7:12"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,2),
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey4))
@
\end{figure}

%% Example of how to create a figure of results
\begin{figure}
\caption{Simulated power curves for Simulation 4. Violation Parameter is covariance}
\label{fig:sim42res}
\setkeys{Gin}{width=\textwidth}
<<sim42res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
print(xyplot(power ~ diff | pars, group = ~ nobs, data = sim4,
             subset = diff <= 4 & parms == "var" & test == "CvM" & pars %in% 
             c("1","7","13","1:6","7:12"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,2),
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey4))
@
\end{figure}

\begin{figure}
\caption{Simulated power curves for Simulation 4. Violation Parameter is error terms}
\label{fig:sim43res}
\setkeys{Gin}{width=\textwidth}
<<sim43res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))

print(xyplot(power ~ diff | pars, group = ~ nobs, data = sim4,
             subset = diff <= 4 & parms == "error" & test == "CvM" & pars %in% 
             c("1","7","13","1:6","7:12"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,2),
             type = "b", xlab="Violation Magnitude", ylab="Power", key=mykey4))
@
\end{figure}

%###########################################################################################

\section{Simulation 6}
Follow sim3. The problem need concern is that when fix diff=3, the plot is similar to sim3. However,
if we look at with diffrent diff in perc=0.5, the plot would be similar to psychometric paper. 
CvM's advantageous is depending on percentage.

%% Example of how to call a simulation in R:
<<sim6>>=
## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1090)

if(file.exists("sim6.rda")) {
  load("sim6.rda")
} else {
  ## Arguments lead to small-scale simulation:
  sim6 <- simulation(nobs=c(100,200,500), diff=seq(0,4,.25), parms=c("loading","var","error"))
  save(sim6, file="sim6.rda")
}

## Add to column names so that plots look nicer
sim6$test <- paste("", sim6$test, sep="")
levels(sim6$perc)<-c(".05",".1",".15",".2",".25",".3",".35",".4",".45",".5",".55",
                     ".6",".65",".7",".75",".8",".85",".9",".95")

@

%% Example of how to create a figure of results
\begin{figure}
\caption{Simulated power curves for Simulation 3.Violation Parameter is loading}
\label{fig:sim61res}
\setkeys{Gin}{width=\textwidth}
<<sim61res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey6 <- simpleKey(unique(sim6$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ perc | pars, group = ~ test, data = sim6,
             subset = diff ==3 & parms == "loading"  & nobs == 500 & pars %in%
             c("1","7","13","1:6","7:12","1:3","7:9") & perc %in% c(".1",".2",
               ".3",".4",".5",".6",".7",".8",".9"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,3),
             type = "b", xlab="Violation Percentage",ylab="Power", key=mykey6))
@
\end{figure}

%% Example of how to create a figure of results
\begin{figure}
\caption{Simulated power curves for Simulation 3.Violation Parameter is loading}
\label{fig:sim62res}
\setkeys{Gin}{width=\textwidth}
<<sim62res, fig=TRUE, height=7, width=7>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey6 <- simpleKey(unique(sim6$test), points = TRUE, lines = TRUE)
print(xyplot(power ~ perc | pars, group = ~ test, data = sim6,
             subset = diff ==3 & parms == "error"  & nobs == 500 & pars %in%
             c("1","7","13","1:6","7:12","1:3","7:9") & perc %in% c(".1",".2",
               ".3",".4",".5",".6",".7",".8",".9"),
              strip = function(which.given, which.panel, var.name,
                        strip.levels = FALSE,
                        strip.names = TRUE, ...) {
       	      strip.default(which.given, which.panel,
       	               var.name = greekpar[which.panel],
                       strip.levels = FALSE,
                       strip.names = TRUE, ...)},
             layout = c(3,3),
             type = "b", xlab="Violation Percentage",ylab="Power", key=mykey6))
@
\end{figure}

\section{General Discussion}

What do the results mean, future studies, etc.

In Simulation 3, the asymmetric pattern is asscociated with parameter value.
Specifically, given unequal groups, when most(percentage is 0.6,0.7,0.8,0.9) of the 
individuals' parameter value is larger, the measurement variance is easier to be detected.
In the situation that violation group parameter value is smaller, as the percentage increases,
individuals' parameter value is smaller, too, the measurement variance is more difficult to be 
detected. Thus the asymmptotic pattern would look reversed (percentage 0.1,0.2,0.3,0.4 would be
associated with larger power compared with the 0.9,0.8,0.7,0.6).

\subsection{Summary}

Concluding paragraph.

\section*{Computational Details}

All results were obtained using the {R}~system for statistical computing \cite{R11},
version~\Sexpr{paste(R.Version()[6:7], collapse = ".")}, employing the add-on package
{lavaan}~\Sexpr{packageDescription("lavaan")["Version"]} \cite{lavaan11} for fitting 
of the factor analysis models and
{strucchange}~\Sexpr{packageDescription("strucchange")["Version"]} \cite{ZeiLei02,Zei06}
for evaluating the parameter instability tests. {R}~and the packages {lavaan} and 
{strucchange} are
freely available under the General Public License~2 from the
Comprehensive {R} Archive Network at \url{http://CRAN.R-project.org/}.
{R}~code for replication of our results is
available at \url{http://semtools.R-Forge.R-project.org/}.

\bibliography{refs}

\end{document}