### 2014-03-04 frontiers/IMPSpres.Rnw
update sim figure part.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@324 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{beamer}
\setbeamertemplate{caption}[numbered]
\usepackage{beamerthemeSzeged,graphics,amsmath,bm,booktabs}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usefonttheme{serif}

\title{Score-based tests of measurement invariance: Use in practice}
\subtitle{2014 Annual Meeting of Psychometric Society}
\author[shortname]{Ting Wang\inst{1} \and Edgar C. Merkle\inst{1} \and 
  Achim Zeileis \inst{2}}
\institute[shortinst]{\inst{1} University of Missouri \and 
                      \inst{2} Universit\"{a}t Innsbruck}

\renewcommand{\baselinestretch}{1.1}

\begin{document}

<<setup, include=FALSE>>=
library("knitr")
opts_chunk$set(size='footnotesize',tidy.opts=list(width.cutoff=50,
               reindent.spaces=6))

## Color unevaluted commands using Sexpr
knit_hooks$set(inline = function(x) { 
  if (is.numeric(x)) return(knitr:::format_sci(x, 'latex'))
  highr:::hi_latex(x)
}) 
## packages
library("lavaan")
library("strucchange") ## at least 1.5-0
library("mvtnorm")
library("lattice")

## auxiliary code
source("../www/sim-frontiers.R")
source("../www/mz-frontiers.R")

## some options
options(prompt = "R> ", continue = "+  ", digits = 3,
  useFancyQuotes = FALSE, show.signif.stars = FALSE)

## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1163)

if(file.exists("sim1.rda")) {
  load("sim1.rda")
} else {
  sim1 <- simulation(sim = "sim1", nobs = 480, parms = c("loading", "error", 
                                                   "var", "intercept"))
  sim1$nlevels <- factor(sim1$nlevels)
  levels(sim1$nlevels) <- paste("m=", levels(sim1$nlevels), sep="")

  sim1$nobs <- factor(sim1$nobs)
  levels(sim1$nobs) <- paste("n=", levels(sim1$nobs), sep="")
  save(sim1, file="sim1.rda")
}
sim1$test <- factor(as.character(sim1$test), levels = c("ordmax","ordwmax",
                                                 "catdiff"),
                    labels = c("maxLM_o", "WDM_o", "LM_uo"))
parlabs <- c(expression(lambda[11]), expression(phi[12]), 
             expression(psi[11]), expression(mu[11]), expression
             (list(lambda[11], ldots, lambda[62])), 
             expression(list(psi[11], ldots, psi[66])), 
             expression(list(mu[11], ldots, mu[66])))
levels(sim1$pars) <- c("Loading1", "Covariance", "Error1", "Intercept1", 
                       "All Loadings", "All Errors", "All Intercepts")  
@

\frame{\titlepage}

\begin{frame}
\frametitle{Problem Scenario}

\end{frame}
%--------------------------------------------------------------------------

\begin{frame}
\frametitle{Overview} 
% Table of contents slide, comment this block out to remove it
\tableofcontents 
% Throughout your presentation, if you choose to use \section{} and 
%\subsection{} commands, these will automatically be printed on this slide 
%as an overview of your presentation
\end{frame}

\section{Theoretical Background}
\section{Simulation: Locating M.I. Violation Parameter Property}
\subsection{Violation Parameter is IN the Model}

\begin{frame}
\begin{figure}
\tiny{\caption{M.I.violation parameter $\lambda_{11}$}}
\label{fig:sim11res}
<<sim11res, fig=TRUE, height=5.5, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
       subset = (parms == "loading" & 
          pars %in% c("Loading1", "Error1", "All Loadings") & diff %% 0.5 == 0),
       type = "b", xlab = 
          expression(paste("Violation Magnitude (", lambda[11], ")")),
       ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
           if(which.given == 2){
             strip.default(which.given, factor.levels = 
                           parlabs[c(1, 3, 5)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

\end{figure}
\end{frame}

\subsection{Violation Parameter is OUT the Model}
\section{Conclusion from Simulation Studies}
\section{Application}

\end{document}