### 2014-03-04 frontiers/IMPSpres.Rnw
update presentation. need to work on application part.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@325 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{beamer}
\setbeamertemplate{caption}[numbered]
\usepackage{beamerthemeSzeged,graphics,amsmath,bm,booktabs}
\usepackage{tikz}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usefonttheme{serif}
% Seems necessary for texlive 2012?
\usetikzlibrary{arrows}

\title{Score-based tests of measurement invariance: Use in practice}
\subtitle{2014 Annual Meeting of Psychometric Society}
\author[shortname]{Ting Wang\inst{1} \and Edgar C. Merkle\inst{1} \and 
  Achim Zeileis \inst{2}}
\institute[]{\inst{1} University of Missouri \and 
                      \inst{2} Universit\"{a}t Innsbruck}

\renewcommand{\baselinestretch}{1.1}

\begin{document}

<<setup, include=FALSE>>=
library("knitr")
opts_chunk$set(size='footnotesize',tidy.opts=list(width.cutoff=45,
               reindent.spaces=6))

## Color unevaluted commands using Sexpr
knit_hooks$set(inline = function(x) { 
  if (is.numeric(x)) return(knitr:::format_sci(x, 'latex'))
  highr:::hi_latex(x)
}) 
## packages
library("lavaan")
library("strucchange") ## at least 1.5-0
library("mvtnorm")
library("lattice")

## auxiliary code
source("../www/sim-frontiers.R")
source("../www/mz-frontiers.R")

## some options
options(prompt = "R> ", continue = "+  ", digits = 3,
  useFancyQuotes = FALSE, show.signif.stars = FALSE)

## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1163)

if(file.exists("sim1.rda")) {
  load("sim1.rda")
} else {
  sim1 <- simulation(sim = "sim1", nobs = 480, parms = c("loading", "error", 
                                                   "var", "intercept"))
  sim1$nlevels <- factor(sim1$nlevels)
  levels(sim1$nlevels) <- paste("m=", levels(sim1$nlevels), sep="")

  sim1$nobs <- factor(sim1$nobs)
  levels(sim1$nobs) <- paste("n=", levels(sim1$nobs), sep="")
  save(sim1, file="sim1.rda")
}
sim1$test <- factor(as.character(sim1$test), levels = c("ordmax","ordwmax",
                                                 "catdiff"),
                    labels = c("maxLM_o", "WDM_o", "LM_uo"))
parlabs <- c(expression(lambda[11]), expression(phi[12]), 
             expression(psi[11]), expression(mu[11]), expression
             (list(lambda[11], ldots, lambda[62])), 
             expression(list(psi[11], ldots, psi[66])), 
             expression(list(mu[11], ldots, mu[66])))
levels(sim1$pars) <- c("Loading1", "Covariance", "Error1", "Intercept1", 
                       "All Loadings", "All Errors", "All Intercepts")

## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1163)

if(file.exists("sim2.rda")) {
  load("sim2.rda")
} else {
  sim2 <- simulation(sim = "sim2", nobs = 480, 
  parms = c("extra", "extra+loading", "extra+var", "extra+error", 
           "extra+intercept"))

  save(sim2, file="sim2.rda")
}
sim2$test <- factor(as.character(sim2$test),
  levels = c("ordmax", "ordwmax", "catdiff"),
  labels = c("maxLM_o", "WDM_o", "LM_uo"))
  sim2$nlevels <- factor(sim2$nlevels)
  levels(sim2$nlevels) <- paste("m=", levels(sim2$nlevels), sep="")
  sim2$nobs <- factor(sim2$nobs)
  levels(sim2$nobs) <- paste("n=", levels(sim2$nobs), sep="")
  parlabs <- c(expression(lambda[11]), expression(phi[12]), 
               expression(psi[11]), expression(mu[11]),
	       expression(list(lambda[11], ldots, lambda[62])), 
               expression(list(psi[11], ldots, psi[66])),
               expression(list(mu[11], ldots, mu[66])))
  levels(sim2$pars) <- c("Loading1", "Covariance", "Error1", "Intercept1", 
                         "All Loadings", "All Errors", "All Intercepts")
@

\frame{\titlepage}

\begin{frame}
\frametitle{Problem Scenario}

\end{frame}
%--------------------------------------------------------------------------

\begin{frame}
\frametitle{Overview} 
% Table of contents slide, comment this block out to remove it
\tableofcontents 
% Throughout your presentation, if you choose to use \section{} and 
%\subsection{} commands, these will automatically be printed on this slide 
%as an overview of your presentation
\end{frame}
%--------------------------------------------------------------------------
\section{Theoretical Background}
\section{Simulation}
\subsection{Violation Parameter is IN the Model}
\begin{frame}
 \begin{figure}
    \caption{Model used for the simulation.}
    \label{fig:pathinmodel}
\resizebox{8.0cm}{!}{
\begin{tikzpicture}
[latent/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=20mm, minimum height=10mm},
 manifest/.style={rectangle,draw=black!50,fill=white!20,thick,
                inner sep=0pt,minimum width=20mm, minimum height=10mm},
 error/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=10mm},
 post/.style={->,shorten >=1pt,>=stealth',semithick},
 postred/.style={->,shorten >=1pt,>=stealth',semithick,red},
 postdashed/.style={->,shorten >=1pt,>=stealth',semithick,red,dashed},
 sling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=20mm,semithick},
 errorsling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=10mm,semithick}]

%position objects.

\node[latent] (VERBAL) at (0,4) {VERBAL};
\node[latent] (MATH)   at (0,0) {MATH};

\node[manifest] (SCALE1) at(5,6) {SCALE1};
\node[manifest] (SCALE2) at (5,4.5) {SCALE2};
\node[manifest] (SCALE3) at (5,3) {SCALE3};
\node[manifest] (SCALE4) at(5,1) {SCALE4};
\node[manifest] (SCALE5) at (5,-0.5) {SCALE5};
\node[manifest] (SCALE6) at (5,-2) {SCALE6};

\node[error] (E1) at (7,6) {E1};
\node[error] (E2) at (7,4.5) {E2};
\node[error] (E3) at (7,3) {E3};
\node[error] (E4) at (7,1) {E4};
\node[error] (E5) at (7,-0.5) {E5};
\node[error] (E6) at (7,-2) {E6};

%arrows, slings and texts
\node[latent] (VERBAL) at (0,4) {VERBAL}
   edge [postred] node[above=0.01cm,text width=2cm]
   {\scriptsize{$\lambda_{11}=4.92$}} (SCALE1)
   edge [post] node[above=0.001mm,text width=2cm]
   {\scriptsize{$\lambda_{21}=2.96$}} (SCALE2)
   edge [post] node[above=0.1cm,text width=2cm]
   {\scriptsize{$\lambda_{31}=5.96$}} (SCALE3)
   edge [sling,bend right=100] node[auto,swap]{1}(VERBAL.-180)
   edge [sling,bend right=70] node[auto,swap]
   {$\phi_{12}=-0.48$}(MATH)  ;

\node[latent] (MATH) at (0,0) {MATH}
   edge [post] node[above=0.01cm,text width=2cm]
   {\scriptsize{$\lambda_{42}=3.24$}}(SCALE4)
   edge [post] node[above=0.01cm,text width=2cm]
   {\scriptsize{$\lambda_{52}=4.32$}}(SCALE5)
   edge [post] node[above=0.25cm,text width=2cm]
   {\scriptsize{$\lambda_{62}=7.21$}}(SCALE6)
   edge [sling,bend right=100] node[auto,swap]{1}(MATH.270);

\node[error] (E1) at (7,6) {E1}
   edge [post] node[auto,swap]{\small{1}}(SCALE1)
   edge [errorsling, bend right=100]node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{11}=26.77$}} (E1.45);
\node[error] (E2) at (7,4.5) {E2}
   edge [post] node[auto,swap]{\small{1}} (SCALE2)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{22}=13.01$}}(E2.45);
\node[error] (E3) at (7,3) {E3}
   edge [post] node[auto,swap]{\small{1}}(SCALE3)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{33}=30.93$}}(E3.45);
\node[error] (E4) at (7,1) {E4}
   edge [post] node[auto,swap]{\small{1}}(SCALE4)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{44}=3.17$}}(E4.45);
\node[error] (E5) at (7,-0.5) {E5}
   edge [post] node[auto,swap]{\small{1}}(SCALE5)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{55}=8.82$}} (E5.45);
\node[error] (E6) at (7,-2) {E6}
   edge [post] node[auto,swap]{\small{1}}(SCALE6)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{66}=22.5$}}(E6.45);
\end{tikzpicture}}
\end{figure}
\end{frame}

\begin{frame}
\begin{figure}
\tiny{\caption{M.I.violation parameter $\lambda_{11}$.  Testing
    parameters are $\lambda_{11}$ and $\lambda_{11} \ldots \lambda_{62}$.}}
\label{fig:sim11res}
<<sim11res, fig=TRUE, height=5, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
       subset = (parms == "loading" & 
          pars %in% c("Loading1", "All Loadings") & diff %% 0.5 == 0),
       type = "b", xlab = 
          expression(paste("Violation Magnitude (", lambda[11], ")")),
       ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
           if(which.given == 2){
             strip.default(which.given, factor.levels = 
                           parlabs[c(1, 5)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

\end{figure}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
\begin{figure}
\tiny{\caption{M.I.violation parameter $\lambda_{11}$.  Testing
    parmaeter is $\psi_{11} \ldots \psi_{66}, \sigma_{12}$ and
    $\mu_{11} \ldots \mu_{66}$.}}
\label{fig:sim12res}
<<sim12res, fig=TRUE, height=5, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
       subset = (parms == "loading" & 
          pars %in% c("All Errors", "Covariance", "All Intercepts") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = 
          expression(paste("Violation Magnitude (", lambda[11], ")")),
       ylab = "Power", ylim = c(0,1), key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
           if(which.given == 2){
             strip.default(which.given, factor.levels = 
                           parlabs[c(2, 6, 7)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

\end{figure}
\end{frame}
%--------------------------------------------------------------------------

\subsection{Violation Parameter is OUT the Model}
\begin{frame}
\begin{figure}
    \caption{Model used for the simulation.}
    \label{fig:pathinmodel}
\resizebox{8.0cm}{!}{
\begin{tikzpicture}
[latent/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=20mm, minimum height=10mm},
 manifest/.style={rectangle,draw=black!50,fill=white!20,thick,
                inner sep=0pt,minimum width=20mm, minimum height=10mm},
 error/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=10mm},
 post/.style={->,shorten >=1pt,>=stealth',semithick},
 postred/.style={->,shorten >=1pt,>=stealth',semithick,red},
 postdashed/.style={->,shorten >=1pt,>=stealth',semithick,red,dashed},
 sling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=20mm,semithick},
 errorsling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=10mm,semithick}]

%position objects.

\node[latent] (VERBAL) at (0,4) {VERBAL};
\node[latent] (MATH)   at (0,0) {MATH};

\node[manifest] (SCALE1) at(5,6) {SCALE1};
\node[manifest] (SCALE2) at (5,4.5) {SCALE2};
\node[manifest] (SCALE3) at (5,3) {SCALE3};
\node[manifest] (SCALE4) at(5,1) {SCALE4};
\node[manifest] (SCALE5) at (5,-0.5) {SCALE5};
\node[manifest] (SCALE6) at (5,-2) {SCALE6};

\node[error] (E1) at (7,6) {E1};
\node[error] (E2) at (7,4.5) {E2};
\node[error] (E3) at (7,3) {E3};
\node[error] (E4) at (7,1) {E4};
\node[error] (E5) at (7,-0.5) {E5};
\node[error] (E6) at (7,-2) {E6};

%arrows, slings and texts
\node[latent] (VERBAL) at (0,4) {VERBAL}
   edge [post] node[above=0.01cm,text width=2cm]
   {\scriptsize{$\lambda_{11}=4.92$}} (SCALE1)
   edge [post] node[above=0.001mm,text width=2cm]
   {\scriptsize{$\lambda_{21}=2.96$}} (SCALE2)
   edge [post] node[above=0.1cm,text width=2cm]
   {\scriptsize{$\lambda_{31}=5.96$}} (SCALE3)
   edge [sling,bend right=100] node[auto,swap]{1}(VERBAL.-180)
   edge [sling,bend right=70] node[auto,swap]
   {$\phi_{12}=-0.48$}(MATH)  ;

\node[latent] (MATH) at (0,0) {MATH}
   edge [post] node[above=0.01cm,text width=2cm]
   {\scriptsize{$\lambda_{42}=3.24$}}(SCALE4)
   edge [postdashed] node [below=0.7cm,text width=2cm, left=0.001cm]
   {\scriptsize{unmodeled loading=$3$}}(SCALE1.210)
   edge [post] node[above=0.01cm,text width=2cm]
   {\scriptsize{$\lambda_{52}=4.32$}}(SCALE5)
   edge [post] node[above=0.25cm,text width=2cm]
   {\scriptsize{$\lambda_{62}=7.21$}}(SCALE6)
   edge [sling,bend right=100] node[auto,swap]{1}(MATH.270);

\node[error] (E1) at (7,6) {E1}
   edge [post] node[auto,swap]{\small{1}}(SCALE1)
   edge [errorsling, bend right=100]node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{11}=26.77$}} (E1.45);
\node[error] (E2) at (7,4.5) {E2}
   edge [post] node[auto,swap]{\small{1}} (SCALE2)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{22}=13.01$}}(E2.45);
\node[error] (E3) at (7,3) {E3}
   edge [post] node[auto,swap]{\small{1}}(SCALE3)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{33}=30.93$}}(E3.45);
\node[error] (E4) at (7,1) {E4}
   edge [post] node[auto,swap]{\small{1}}(SCALE4)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{44}=3.17$}}(E4.45);
\node[error] (E5) at (7,-0.5) {E5}
   edge [post] node[auto,swap]{\small{1}}(SCALE5)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{55}=8.82$}} (E5.45);
\node[error] (E6) at (7,-2) {E6}
   edge [post] node[auto,swap]{\small{1}}(SCALE6)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{66}=22.5$}}(E6.45);
\end{tikzpicture}}
\end{figure}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
\begin{figure}
\tiny{\caption{The parameter violating measurement invariance is the unmodeled
loading. Testing parameters are $\lambda_{11}$ and $\psi_{11}$.}}
\label{fig:sim21res}
<<sim21res, fig=TRUE, height=5, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra" & 
	 pars %in% c("Loading1", "Error1") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude (Unmodeled Loading)",
       ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	 strip.default(which.given, factor.levels = parlabs[c(1, 3)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@
\end{figure}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
\begin{figure}
\tiny{\caption{The parameter violating measurement invariance is the unmodeled
loading. Testing parameters are $\sigma_{12}$ and $\mu_{11}$.}}
\label{fig:sim22res}
<<sim22res, fig=TRUE, height=5, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra" & 
	 pars %in% c("Covariance", "Intercept1") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude (Unmodeled Loading)",
       ylab = "Power", ylim = c(0,1), key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	 strip.default(which.given, factor.levels = parlabs[c(2, 4)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@
\end{figure}
\end{frame}

\section{Conclusion}
\section{Application}
\begin{frame}[fragile]
\frametitle{Input data and Fit model}
\begin{enumerate}
 \item Input data
<<data,eval=FALSE>>=
data("YouthGratitude", package = "psychotools")
compcases <- apply(YouthGratitude[, 4:28], 1, function(x) all(x %in% 1:9))
yg <- YouthGratitude[compcases, ]
@ 
\item Fit restricted model and full model
<<modeest,eval=FALSE>>=
restr <- cfa("f1 =~ gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
  data = yg, group = "agegroup", meanstructure = TRUE,
  group.equal = "loadings")

full <- cfa("f1 =~ gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
  data = yg, group = "agegroup", meanstructure = TRUE)
@ 
\end{enumerate}
\end{frame}

%--------------------------------------------------------------------------
%TODO: reset counter for enumerate.
\begin{frame}
  \item fff
\end{frame}
\end{document}