### 2014-03-13 frontiers/IMPSpres.Rnw
update theory part, need to work on minor details.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@328 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass{beamer}
\setbeamertemplate{caption}[numbered]
\usepackage{beamerthemeSzeged,graphics,amsmath,bm,booktabs}
\usepackage{tikz}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usefonttheme{serif}
% Seems necessary for texlive 2012?
\usetikzlibrary{arrows}

\title{Score-based tests of measurement invariance: Use in practice}
\subtitle{2014 Annual Meeting of Psychometric Society}
\author[shortname]{Ting Wang\inst{1} \and Edgar C. Merkle\inst{1} \and 
  Achim Zeileis \inst{2}}
\institute[]{\inst{1} University of Missouri \and 
                      \inst{2} Universit\"{a}t Innsbruck}

\renewcommand{\baselinestretch}{1.1}
\newcounter{sauvegardeenumi}
\newcommand{\asuivre}{\setcounter{sauvegardeenumi}{\theenumi}}
\newcommand{\suite}{\setcounter{enumi}{\thesauvegardeenumi}}

\begin{document}

<<setup, include=FALSE>>=
library("knitr")
opts_chunk$set(size='footnotesize',tidy.opts=list(width.cutoff=45,
               reindent.spaces=6))

## Color unevaluted commands using Sexpr
knit_hooks$set(inline = function(x) { 
  if (is.numeric(x)) return(knitr:::format_sci(x, 'latex'))
  highr:::hi_latex(x)
}) 
## packages
library("lavaan")
library("strucchange") ## at least 1.5-0
library("mvtnorm")
library("lattice")

## auxiliary code
source("../www/sim-frontiers.R")
source("../www/mz-frontiers.R")

## some options
options(prompt = "R> ", continue = "+  ", digits = 3,
  useFancyQuotes = FALSE, show.signif.stars = FALSE)

## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1163)

if(file.exists("sim1.rda")) {
  load("sim1.rda")
} else {
  sim1 <- simulation(sim = "sim1", nobs = 480, parms = c("loading", "error", 
                                                   "var", "intercept"))
  sim1$nlevels <- factor(sim1$nlevels)
  levels(sim1$nlevels) <- paste("m=", levels(sim1$nlevels), sep="")

  sim1$nobs <- factor(sim1$nobs)
  levels(sim1$nobs) <- paste("n=", levels(sim1$nobs), sep="")
  save(sim1, file="sim1.rda")
}
sim1$test <- factor(as.character(sim1$test), levels = c("ordmax","ordwmax",
                                                 "catdiff"),
                    labels = c("maxLM_o", "WDM_o", "LM_uo"))
parlabs <- c(expression(lambda[11]), expression(phi[12]), 
             expression(psi[11]), expression(mu[11]), expression
             (list(lambda[11], ldots, lambda[62])), 
             expression(list(psi[11], ldots, psi[66])), 
             expression(list(mu[11], ldots, mu[66])))
levels(sim1$pars) <- c("Loading1", "Covariance", "Error1", "Intercept1", 
                       "All Loadings", "All Errors", "All Intercepts")

## seed for replication
RNGkind(kind = "default", normal.kind = "default")
set.seed(1163)

if(file.exists("sim2.rda")) {
  load("sim2.rda")
} else {
  sim2 <- simulation(sim = "sim2", nobs = 480, 
  parms = c("extra", "extra+loading", "extra+var", "extra+error", 
           "extra+intercept"))

  save(sim2, file="sim2.rda")
}
sim2$test <- factor(as.character(sim2$test),
  levels = c("ordmax", "ordwmax", "catdiff"),
  labels = c("maxLM_o", "WDM_o", "LM_uo"))
  sim2$nlevels <- factor(sim2$nlevels)
  levels(sim2$nlevels) <- paste("m=", levels(sim2$nlevels), sep="")
  sim2$nobs <- factor(sim2$nobs)
  levels(sim2$nobs) <- paste("n=", levels(sim2$nobs), sep="")
  parlabs <- c(expression(lambda[11]), expression(phi[12]), 
               expression(psi[11]), expression(mu[11]),
	       expression(list(lambda[11], ldots, lambda[62])), 
               expression(list(psi[11], ldots, psi[66])),
               expression(list(mu[11], ldots, mu[66])))
  levels(sim2$pars) <- c("Loading1", "Covariance", "Error1", "Intercept1", 
                         "All Loadings", "All Errors", "All Intercepts")

data("YouthGratitude", package = "psychotools")
compcases <- apply(YouthGratitude[, 4:28], 1, function(x) all(x %in% 1:9))
yg <- YouthGratitude[compcases, ]

restr <- cfa("f1 =~ gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
  data = yg, group = "agegroup", meanstructure = TRUE,
  group.equal = "loadings")

set.seed(1090)
wdmo <- sctest(restr, order.by = yg$agegroup, parm = 1:4,
  vcov = "info", functional = "WDMo")

if(file.exists("maxLMo.rda")){
  load("maxLMo.rda")
} else {
  RNGkind(kind = "default", normal.kind = "default")
  set.seed(1090)
  maxLMo <- ordL2BB(yg$agegroup, nproc = 4)
  save(maxLMo, file = "maxLMo.rda")
}
maxlmo <- sctest(restr, order.by = yg$agegroup, parm = 1:4,
  vcov = "info", functional = maxLMo)

@

\frame{\titlepage}

\begin{frame}
\frametitle{Problem Scenario}
\begin{itemize}
  \item
 \emph{Data:} The data consist of responses to a gratitude scale from $n=1401$ youth 
 aged 10--19 years, dividing into six age groups: 10-11, 12-13, 14,
 15, 16, 17-19. 
  \item
\emph{Question:} The researcher is interested in whether the scales 
were measurement invariant across these age groups.  
\end{itemize}

\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{Similar Problem Scenarios and Generalization}
  \begin{itemize}
    \item Whether model parameter holds for people with different
      income level, grade level, motivation level, etc.  
    \item In general, the problem is whether measurement invariance
      hypothesis holds with an auxilliary ordinal variable. 
    \end{itemize}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
\frametitle{Possible Solution}
\begin{itemize}
\item Likelihood Ratio Test(LRT) for full model and reduced model(s).
  \begin{itemize}
    \item Ignore the nature of ordinal variable, treat it as
      categorical varaible. 
    \item Sensitive to sample size. Rejecte the reduced model(s) too often.
    \item Rather tedious and time consuming to fit and compare all possible
      models. e.g, for testing one parameter, you need to fit $*2^{m-1}$ models,
      $m$ is the level of ordinal variable.
    \end{itemize}
\item Score-based Tests w.r.t. ordinal variable. 
\end{itemize}
  \end{frame}
\begin{frame}
\frametitle{Overview} 
% Table of contents slide, comment this block out to remove it
\tableofcontents 
% Throughout your presentation, if you choose to use \section{} and 
%\subsection{} commands, these will automatically be printed on this slide 
%as an overview of your presentation
\end{frame}
%--------------------------------------------------------------------------
\section{Theoretical Background}
\begin{frame}
  \frametitle{Score for Each Observation}
\begin{equation}
\label{eq:score}
s(\theta;x_{i}) = \left (
  \frac{\partial \ell(\theta;x_{i})}{\partial \theta_{1}}, \dots,
  \frac{\partial \ell(\theta;x_{i})}{\partial \theta_{k}} \right
  )^\top,\ i=1,\ldots,n,
\end{equation} 
\scriptsize{where $\ell(\theta; x_i)$ is the likelihood associated with individual
$i$ and $\theta$ is a $k$-dimensional parameter vector. The corresponding
maximum likelihood estimate $\hat{\theta}$ solves the first order condition:
$\sum_{i = 1}^n s(\hat \theta;x_{i}) = 0$.}  
\end{frame}
%--------------------------------------------------------------------------

\begin{frame}
  \frametitle{Decorrelate and Order Cumulative Score}
To formalize these ideas, we define a suitably scaled cumulative sum of the 
ordered scores.  This can be written as
\begin{equation} 
    \label{eq:cumscore}
  {\bm B}(t; \hat {\bm \theta}) ~=~ \hat {\bm I}^{-1/2} n^{-1/2}
    \sum_{i = 1}^{\lfloor n \cdot t \rfloor} {\bm s}(\hat {\bm \theta}; x_{(i)})
  \qquad (0 \le t \le 1)
\end{equation}
\scriptsize{where $\hat{\bm I}$ is an estimate of the information matrix, $\lfloor
nt \rfloor$ is the integer part of $nt$, and $x_{(i)}$ reflects the individual 
with the $i$-thsmallest value of the auxiliary variable $V$.} 
  \end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{}
  \begin{itemize}
   \item  Under the hypothesis of measurement invariance, a central limit theorem can be
   used to show that the fluctuation of the above cumulative sum follows a
   Brownian bridge.
   \item This result allows us to calculate $p$-values and
   critical values for test statistics under the hypothesis of
   measurement invariance.
  \end{itemize}
 \end{frame}
%--------------------------------------------------------------------------  
 \begin{frame}
   \frametitle{Statistics Based on Decorrelated and Ordered Socre}
   Multiple test statistics are available, depending on how one
summarizes the behavior of the cumulative sum of scores. Below shows
 two statistics for ordinal auxiliary variable with $m$ levels,

\scriptsize{\begin{eqnarray}
    \label{eq:wdm}
      \mathit{WDM}_o & = & \max_{i \in \{i_1, \dots, i_{m-1} \}} ~
      \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1/2}      
                             \max_{j = 1, \dots, k} | {\bm B}(\hat {\bm \theta})_{ij} |\\
    \label{eq:maxlmo}
      \max \mathit{LM}_o & = & \max_{i \in \{i_1, \dots, i_{m-1} \}} ~
                             \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1}
                             \sum_{j = 1, \dots, k} {\bm B}(\hat {\bm \theta})_{ij}^2
\end{eqnarray}}
where $i_{\ell}=\lfloor n \cdot t_{\ell} \rfloor$ $(\ell=1,\ldots,m-1)$.
   \end{frame}
%--------------------------------------------------------------------------  
\begin{frame}
  \frametitle{Visual Explanation of the Score-based Test for ordinal
    auxillary varaible.}
$$\bordermatrix{ \text{$B$}& \theta_{1} & \theta_{2} & \ldots &\theta_{k}\cr
                x_{1}&  1.32  & -4.32 & \ldots & -2.34\cr
                x_{1}&  3.37 & 8.32 & \ldots & \textcolor{red}{63.45}\cr
                x_{..}& \vdots & \vdots & \vdots & \vdots\cr
                x_{m}&  0 & 0 & \ldots & 0\cr
                x_{m}& 0     &   0     & \ldots & 0}$$

  \end{frame}

\section{Simulation}
\subsection{Violation Parameter is IN the Model}
\begin{frame}
 \begin{figure}
    \caption{Model used for the simulation.}
    \label{fig:pathinmodel}
\resizebox{8.0cm}{!}{
\begin{tikzpicture}
[latent/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=20mm, minimum height=10mm},
 manifest/.style={rectangle,draw=black!50,fill=white!20,thick,
                inner sep=0pt,minimum width=20mm, minimum height=10mm},
 error/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=10mm},
 post/.style={->,shorten >=1pt,>=stealth',semithick},
 postred/.style={->,shorten >=1pt,>=stealth',semithick,red},
 postdashed/.style={->,shorten >=1pt,>=stealth',semithick,red,dashed},
 sling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=20mm,semithick},
 errorsling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=10mm,semithick}]

%position objects.

\node[latent] (VERBAL) at (0,4) {VERBAL};
\node[latent] (MATH)   at (0,0) {MATH};

\node[manifest] (SCALE1) at(5,6) {SCALE1};
\node[manifest] (SCALE2) at (5,4.5) {SCALE2};
\node[manifest] (SCALE3) at (5,3) {SCALE3};
\node[manifest] (SCALE4) at(5,1) {SCALE4};
\node[manifest] (SCALE5) at (5,-0.5) {SCALE5};
\node[manifest] (SCALE6) at (5,-2) {SCALE6};

\node[error] (E1) at (7,6) {E1};
\node[error] (E2) at (7,4.5) {E2};
\node[error] (E3) at (7,3) {E3};
\node[error] (E4) at (7,1) {E4};
\node[error] (E5) at (7,-0.5) {E5};
\node[error] (E6) at (7,-2) {E6};

%arrows, slings and texts
\node[latent] (VERBAL) at (0,4) {VERBAL}
   edge [postred] node[above=0.01cm,text width=2cm]
   {\scriptsize{$\lambda_{11}=4.92$}} (SCALE1)
   edge [post] node[above=0.001mm,text width=2cm]
   {\scriptsize{$\lambda_{21}=2.96$}} (SCALE2)
   edge [post] node[above=0.1cm,text width=2cm]
   {\scriptsize{$\lambda_{31}=5.96$}} (SCALE3)
   edge [sling,bend right=100] node[auto,swap]{1}(VERBAL.-180)
   edge [sling,bend right=70] node[auto,swap]
   {$\phi_{12}=-0.48$}(MATH)  ;

\node[latent] (MATH) at (0,0) {MATH}
   edge [post] node[above=0.01cm,text width=2cm]
   {\scriptsize{$\lambda_{42}=3.24$}}(SCALE4)
   edge [post] node[above=0.01cm,text width=2cm]
   {\scriptsize{$\lambda_{52}=4.32$}}(SCALE5)
   edge [post] node[above=0.25cm,text width=2cm]
   {\scriptsize{$\lambda_{62}=7.21$}}(SCALE6)
   edge [sling,bend right=100] node[auto,swap]{1}(MATH.270);

\node[error] (E1) at (7,6) {E1}
   edge [post] node[auto,swap]{\small{1}}(SCALE1)
   edge [errorsling, bend right=100]node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{11}=26.77$}} (E1.45);
\node[error] (E2) at (7,4.5) {E2}
   edge [post] node[auto,swap]{\small{1}} (SCALE2)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{22}=13.01$}}(E2.45);
\node[error] (E3) at (7,3) {E3}
   edge [post] node[auto,swap]{\small{1}}(SCALE3)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{33}=30.93$}}(E3.45);
\node[error] (E4) at (7,1) {E4}
   edge [post] node[auto,swap]{\small{1}}(SCALE4)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{44}=3.17$}}(E4.45);
\node[error] (E5) at (7,-0.5) {E5}
   edge [post] node[auto,swap]{\small{1}}(SCALE5)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{55}=8.82$}} (E5.45);
\node[error] (E6) at (7,-2) {E6}
   edge [post] node[auto,swap]{\small{1}}(SCALE6)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{66}=22.5$}}(E6.45);
\end{tikzpicture}}
\end{figure}
\end{frame}

\begin{frame}
\begin{figure}
\tiny{\caption{M.I.violation parameter $\lambda_{11}$.  Testing
    parameters are $\lambda_{11}$ and $\lambda_{11} \ldots \lambda_{62}$.}}
\label{fig:sim11res}
<<sim11res, fig=TRUE, height=5, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
       subset = (parms == "loading" & 
          pars %in% c("Loading1", "All Loadings") & diff %% 0.5 == 0),
       type = "b", xlab = 
          expression(paste("Violation Magnitude (", lambda[11], ")")),
       ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
           if(which.given == 2){
             strip.default(which.given, factor.levels = 
                           parlabs[c(1, 5)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

\end{figure}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
\begin{figure}
\tiny{\caption{M.I.violation parameter $\lambda_{11}$.  Testing
    parmaeter is $\psi_{11} \ldots \psi_{66}, \sigma_{12}$ and
    $\mu_{11} \ldots \mu_{66}$.}}
\label{fig:sim12res}
<<sim12res, fig=TRUE, height=5, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim1$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim1,
       subset = (parms == "loading" & 
          pars %in% c("All Errors", "Covariance", "All Intercepts") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = 
          expression(paste("Violation Magnitude (", lambda[11], ")")),
       ylab = "Power", ylim = c(0,1), key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
           if(which.given == 2){
             strip.default(which.given, factor.levels = 
                           parlabs[c(2, 6, 7)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@

\end{figure}
\end{frame}
%--------------------------------------------------------------------------

\subsection{Violation Parameter is OUT the Model}
\begin{frame}
\begin{figure}
    \caption{Model used for the simulation.}
    \label{fig:pathinmodel}
\resizebox{8.0cm}{!}{
\begin{tikzpicture}
[latent/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=20mm, minimum height=10mm},
 manifest/.style={rectangle,draw=black!50,fill=white!20,thick,
                inner sep=0pt,minimum width=20mm, minimum height=10mm},
 error/.style={circle,draw=black!50,fill=white!20,thick,
                inner sep=0pt, minimum width=10mm},
 post/.style={->,shorten >=1pt,>=stealth',semithick},
 postred/.style={->,shorten >=1pt,>=stealth',semithick,red},
 postdashed/.style={->,shorten >=1pt,>=stealth',semithick,red,dashed},
 sling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=20mm,semithick},
 errorsling/.style={<->,shorten >=1pt,>=stealth',shorten <=1pt,>=stealth',
   auto,distance=10mm,semithick}]

%position objects.

\node[latent] (VERBAL) at (0,4) {VERBAL};
\node[latent] (MATH)   at (0,0) {MATH};

\node[manifest] (SCALE1) at(5,6) {SCALE1};
\node[manifest] (SCALE2) at (5,4.5) {SCALE2};
\node[manifest] (SCALE3) at (5,3) {SCALE3};
\node[manifest] (SCALE4) at(5,1) {SCALE4};
\node[manifest] (SCALE5) at (5,-0.5) {SCALE5};
\node[manifest] (SCALE6) at (5,-2) {SCALE6};

\node[error] (E1) at (7,6) {E1};
\node[error] (E2) at (7,4.5) {E2};
\node[error] (E3) at (7,3) {E3};
\node[error] (E4) at (7,1) {E4};
\node[error] (E5) at (7,-0.5) {E5};
\node[error] (E6) at (7,-2) {E6};

%arrows, slings and texts
\node[latent] (VERBAL) at (0,4) {VERBAL}
   edge [post] node[above=0.01cm,text width=2cm]
   {\scriptsize{$\lambda_{11}=4.92$}} (SCALE1)
   edge [post] node[above=0.001mm,text width=2cm]
   {\scriptsize{$\lambda_{21}=2.96$}} (SCALE2)
   edge [post] node[above=0.1cm,text width=2cm]
   {\scriptsize{$\lambda_{31}=5.96$}} (SCALE3)
   edge [sling,bend right=100] node[auto,swap]{1}(VERBAL.-180)
   edge [sling,bend right=70] node[auto,swap]
   {$\phi_{12}=-0.48$}(MATH)  ;

\node[latent] (MATH) at (0,0) {MATH}
   edge [post] node[above=0.01cm,text width=2cm]
   {\scriptsize{$\lambda_{42}=3.24$}}(SCALE4)
   edge [postdashed] node [below=0.7cm,text width=2cm, left=0.001cm]
   {\scriptsize{unmodeled loading=$3$}}(SCALE1.210)
   edge [post] node[above=0.01cm,text width=2cm]
   {\scriptsize{$\lambda_{52}=4.32$}}(SCALE5)
   edge [post] node[above=0.25cm,text width=2cm]
   {\scriptsize{$\lambda_{62}=7.21$}}(SCALE6)
   edge [sling,bend right=100] node[auto,swap]{1}(MATH.270);

\node[error] (E1) at (7,6) {E1}
   edge [post] node[auto,swap]{\small{1}}(SCALE1)
   edge [errorsling, bend right=100]node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{11}=26.77$}} (E1.45);
\node[error] (E2) at (7,4.5) {E2}
   edge [post] node[auto,swap]{\small{1}} (SCALE2)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{22}=13.01$}}(E2.45);
\node[error] (E3) at (7,3) {E3}
   edge [post] node[auto,swap]{\small{1}}(SCALE3)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{33}=30.93$}}(E3.45);
\node[error] (E4) at (7,1) {E4}
   edge [post] node[auto,swap]{\small{1}}(SCALE4)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{44}=3.17$}}(E4.45);
\node[error] (E5) at (7,-0.5) {E5}
   edge [post] node[auto,swap]{\small{1}}(SCALE5)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{55}=8.82$}} (E5.45);
\node[error] (E6) at (7,-2) {E6}
   edge [post] node[auto,swap]{\small{1}}(SCALE6)
   edge [errorsling, bend right=100] node[right=0.1,text width=2cm]
   {\scriptsize{$\psi_{66}=22.5$}}(E6.45);
\end{tikzpicture}}
\end{figure}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
\begin{figure}
\tiny{\caption{The parameter violating measurement invariance is the unmodeled
loading. Testing parameters are $\lambda_{11}$ and $\psi_{11}$.}}
\label{fig:sim21res}
<<sim21res, fig=TRUE, height=5, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra" & 
	 pars %in% c("Loading1", "Error1") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude (Unmodeled Loading)",
       ylab = "Power", key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	 strip.default(which.given, factor.levels = parlabs[c(1, 3)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@
\end{figure}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
\begin{figure}
\tiny{\caption{The parameter violating measurement invariance is the unmodeled
loading. Testing parameters are $\sigma_{12}$ and $\mu_{11}$.}}
\label{fig:sim22res}
<<sim22res, fig=TRUE, height=5, width=9, echo=FALSE>>=
trellis.par.set(theme = canonical.theme(color = FALSE))
mykey <- simpleKey(levels(sim2$test), points = TRUE, lines = TRUE)
xyplot(power ~ diff | nlevels + pars, group = ~ test, data = sim2,
       subset = (parms == "extra" & 
	 pars %in% c("Covariance", "Intercept1") 
                 & diff %% 0.5 == 0),
       type = "b", xlab = "Violation Magnitude (Unmodeled Loading)",
       ylab = "Power", ylim = c(0,1), key = mykey, as.table = TRUE,
       strip = function(..., which.given, factor.levels){
	   if(which.given == 2){
	 strip.default(which.given, factor.levels = parlabs[c(2, 4)], ...)
	   } else {
	       strip.default(which.given, factor.levels = factor.levels, ...)
	   }
       })
@
\end{figure}
\end{frame}

%--------------------------------------------------------------------------
\section{Conclusion}

\begin{frame}
  \frametitle{Conclusion based on Simulation Studies}
 \begin{itemize}
   \item
      The ordinal test statistic proposed have relatively-high power for detecting measurement 
      invariance violation parameter specifically that are monotonic with the auxiliary ordinal variable. 
   \item This specificity feature is robust as long as the violation parameter is 
     included in the testing model.
   \item The use of these tests is convenient without changing the focal psychometric model.
   \end{itemize}
\end{frame}
%--------------------------------------------------------------------------
\section{Application}
\begin{frame}[fragile]
\frametitle{Procedures of Application}
\begin{enumerate}[1.]
 \item Input data
\end{enumerate}
<<data,eval=FALSE>>=
data("YouthGratitude", package = "psychotools")
compcases <- apply(YouthGratitude[, 4:28], 1, function(x) all(x %in% 1:9))
yg <- YouthGratitude[compcases, ]
@

\end{frame}

%--------------------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{Procedures of Application}
\begin{enumerate}[2.]
  \suite
\item Fit your model:no need to comparing various models.
\end{enumerate}
 <<modeest,eval=FALSE>>=
 library(lavaan)
  restr <- cfa("f1 =~ gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
  data = yg, group = "agegroup", meanstructure = TRUE,
  group.equal = "loadings")
@

\end{frame}
%--------------------------------------------------------------------------
\begin{frame}[fragile]
\frametitle{Procedures of Application}
\begin{enumerate}[3.]
  \suite
  \item Test if M.I. is violated.
\end{enumerate}
<<sctest-ord2, eval=FALSE>>=
 library(strucchange)
wdmo <- sctest(restr, order.by = yg$agegroup, parm = 1:4, vcov = "info",
  functional = "WDMo",  plot = TRUE, ylim = c(0, 3.2),
  xlab = "Age group", ylab = "Weighted test statistics")
@ 
\end{frame}
%--------------------------------------------------------------------------

\begin{frame}
\begin{figure}
\tiny{\caption{Fluctuation processes for the $\mathit{WDM}_o$ statistic}}
\label{fig:ordres}
<<sctest-ord-graphics,fig=TRUE, height=3.3, width=5, echo=FALSE, results=hide>>=
sctest(restr, order.by = yg$agegroup, parm = 1:4, vcov = "info",
  functional = "WDMo", plot = TRUE, ylim = c(0, 3.2),
  xlab = "Age group", ylab = "Weighted test statistics",
  main = expression(WDM[o]), boundary = list(col = 1, lty = 2))
@
\end{figure}
\end{frame}

\end{document}