### 2015-07-22 ecpa2015/slides.Rnw
first draft for ECPA13 slides
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@377 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[11pt,compress,t]{beamer}
\usetheme{Z}
\usepackage{amsfonts,amstext,amsmath,bm}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}

%% need no \usepackage{Sweave}
\definecolor{InputColor}{rgb}{0,0,0.3}
\definecolor{OutputColor}{rgb}{0.2,0.2,0.2}

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
set.seed(1090)
@

\begin{document}

\title{{\fontsize{14.7}{16}\selectfont Score-Based Tests of Measurement Invariance with Respect to Continuous and Ordinal Variables}}
\author{Achim Zeileis, Edgar C. Merkle, Ting Wang}
\URL{http://eeecon.uibk.ac.at/~zeileis/}
\lecture{Score-Based Tests of Measurement Invariance with Respect to Continuous and Ordinal Variables}

\subsection{Overview}

\begin{frame}
\frametitle{Overview}

\begin{itemize}
  \item Motivation
  \item Framework
  \item Score-based tests
  \begin{itemize}
    \item Continuous variables
    \item Ordinal variables
    \item Categorial variables
  \end{itemize}
  \item Illustration
\end{itemize}

\end{frame}

\subsection{Motivation}

\begin{frame}
\frametitle{Motivation}

\textbf{Psychometric models:} Typically measure latent scales based on certain
manifest variables, e.g., item response theory (IRT) models or confirmatory
factor analysis (CFA, today's focus).

\medskip

\textbf{Crucial assumption:} Measurement invariance (MI). Otherwise observed differences in
scores cannot be reliably attributed to the latent variable that the model
purports to measure.

\medskip

\textbf{Parameter stability:} In parametric models, the MI assumption
corresponds to stability of parameters across all possible subgroups.

\medskip

\textbf{Inference:} The typical approach for assessing MI is
\begin{itemize}
  \item to split the data into reference and focal groups,
  \item assess the stability of selected parameters (all or only a subset) across these groups
  \item by means of standard tests: likelihood ratio (LR), Wald, or Lagrange multiplier
    (LM or score) tests.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Motivation}

\textbf{Problems:}
\begin{itemize}
  \item Subgroups have to be formed in advance.
  \item Numerical variables are often categorized into groups in an ad hoc way
    (e.g., splitting at the median).
  \item In ordinal variables the ordering of the categories is often not exploited
    (assessing only if at least one group differs from the others).
  \item When likelihood ratio or Wald tests are employed, the model has to be
    fitted to each subgroup which can become numerically challenging and computationally
    intensive.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Motivation}

\textbf{Idea:}
\begin{itemize}
  \item Generalize the LM test (or score test).
  \item Thus, the model only has to be fitted once under the MI assumption to the full data set.
  \item Catpure model deviations along a variable that is suspected to cause MI violations.
  \item Exploit ordering to assess if there is (at least) one split so that the
    model parameters before and after the split differ.
  \item The split does \emph{not} have to be known or guessed in advance.
\end{itemize}

\medskip

\textbf{Illustration:} CFA for artificial data.
\begin{itemize}
  \item Model with two latent scales: Verbal and math score.
  \item Three manifest variables for each scale.
  \item Violation of MI for the math loadings along the age of the subjects.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Motivation: CFA for age $\le 16$}

\vspace*{-0.6cm}

\includegraphics[width=0.95\textwidth]{../psychoco2012/model_under16.pdf}

\end{frame}

\begin{frame}
\frametitle{Motivation: CFA for age $> 16$}

\vspace*{-0.6cm}

\includegraphics[width=0.95\textwidth]{../psychoco2012/model_over16.pdf}

\end{frame}

\subsection{Framework}

\begin{frame}
\frametitle{Framework}

\textbf{Model:} Based on log-likelihood $\ell(\cdot)$ for
$p$-dimensional observations $\bm{x}_i$ ($i = 1, \dots, n$)
based on $k$-dimensional parameter ${\bm \theta}$.

\medskip

\textbf{Estimation:} Maximum likelihood.
\[
  \hat{\bm{\theta}} ~=~ \argmax_{\bm{\theta}} \sum_{i = 1}^n \ell(\bm{\theta}; \bm{x}_i).
\]

\medskip

\textbf{Equivalently:} Solve first order conditions
\[
  \sum_{i=1}^{n} {\bm s}(\hat{\bm{\theta}}; \bm{x}_i) ~=~ 0,    
\]
where the score function is the partial derivative of the casewise
likelihood contributions w.r.t.\ the parameters $\bm{\theta}$.
\[
  {\bm s}({\bm \theta}; \bm{x}_i) ~=~ \left(
    \frac{\partial \ell({\bm \theta}; \bm{x}_i)}{\partial \theta_1},
    \dots,
    \frac{\partial \ell({\bm \theta}; \bm{x}_i)}{\partial \theta_k}
  \right)^\top.
\]

\end{frame}

\begin{frame}
\frametitle{Framework}

\textbf{Assumption:} Distribution/likelihood of $\bm{x}_i$ depends
only on the latent scales (through the parameters $\bm{\theta}$) --
but not on any other variable $v_i$.

\medskip

\textbf{Alternative view:} Parameters $\bm{\theta}$ do not depend
any such variable $v_i$. Hence assess for $i = 1, \ldots, n$
\begin{eqnarray*}
    H_0: {\bm \theta}_i & = & {\bm \theta}_0, \\
    H_1: {\bm \theta}_i & = & {\bm \theta}(v_i).
\end{eqnarray*}

\medskip

\textbf{Special case:} Two subgroups resulting from one split point $\nu$.
\begin{eqnarray*}
  H_1^*: {\bm \theta}_i & = & \left\{ \begin{array}{ll}
    {\bm \theta}^{(A)} & \mbox{if } v_i \le \nu \\
    {\bm \theta}^{(B)} & \mbox{if } v_i >   \nu
  \end{array} \right.
\end{eqnarray*}

\medskip

\textbf{Tests:} LR/Wald/LM tests can be easily employed if pattern $\bm{\theta}(v_i)$
is known, specifically for $H_1^*$ with fixed split point $\nu$.

\end{frame}

\begin{frame}
\frametitle{Framework}

\textbf{For unknown split points:} Compute LR/Wald/LM tests for each possible
split point $v_1 \le v_2 \le \dots \le v_n$ and reject if the maximum statistic is large.

\medskip

\textbf{Caution:} By maximally selecting the test statistic different critical
values are required (not from a $\chi^2$ distribution)!

\medskip

\textbf{Illustration:} Assess all $k^* = 19$ model parameters from the artificial
CFA example along the continuous variable age ($v_i$).

\end{frame}

\begin{frame}
\frametitle{Framework}

\vspace*{-1.2cm}

\includegraphics[width=\textwidth]{../psychoco2012/paper-lm-lr-19.pdf}

\end{frame}

\begin{frame}
\frametitle{Framework}

\textbf{Note:} For the maxLM test the parameters $\hat{\bm \theta}$ only have
to be estimated once. Only the model scores ${\bm s}(\hat{\bm \theta}; \bm{x}_i)$
have to be aggregated differently for each split point.

\medskip

\textbf{More generally:} Consider a class of tests that
assesses whether the model ``deviations'' ${\bm s}(\hat{\bm \theta}; \bm{x}_i)$
depend on $v_i$. This can consider only a subset $k*$ of all $k$
parameters/scores or try to capture other patterns than $H_1^*$.

\end{frame}

\begin{frame}
\frametitle{Score-based tests}

\textbf{Fluctuation process:} Capture fluctuations in the cumulative
sum of the scores ordered by the variable $v$.
\[
  {\bm B}(t; \hat {\bm \theta}) ~=~ \hat {\bm I}^{-1/2} n^{-1/2}
    \sum_{i = 1}^{\lfloor n \cdot t \rfloor} {\bm s}(\hat {\bm \theta}; \bm{x}_{(i)})
  \qquad (0 \le t \le 1).
\]
\begin{itemize}
  \item $\hat{\bm I}$ -- estimate of the information matrix.
  \item $t$ -- proportion of data ordered by $v$.
  \item $\lfloor n \cdot t \rfloor$ -- integer part of $n \cdot t$.
  \item $x_{(i)}$ -- observation with the $i$-th smallest value of the
    variable $v$.
\end{itemize}

\medskip

\textbf{Functional central limit theorem:} Under $H_0$ convergence to a (continuous) Brownian bridge process
${\bm B}(\cdot; \hat {\bm \theta}) ~\overset{d}{\rightarrow}~ {\bm B}^{0}(\cdot)$,
from which critical values can be obtained -- either analytically or by simulation.

\end{frame}

\begin{frame}
\frametitle{Score-based tests: Continuous variables}

\textbf{Test statistics:} The empirical process can be viewed as a matrix
${\bm B}(\hat {\bm \theta})_{ij}$ with rows $i = 1, \dots, n$ (observations)
and columns $j = 1, \dots, k$ (parameters). This can be aggregated to scalar
test statistics along continuous the variable $v$.
\begin{eqnarray*}
    \mathit{DM}      & = & \max_{i = 1,\dots, n} \max_{j = 1, \dots, k} | {\bm B}(\hat {\bm \theta})_{ij} | \\
    \mathit{CvM}     & = & n^{-1} \sum_{i = 1,\dots, n} \sum_{j = 1, \dots, k} {\bm B}(\hat {\bm \theta})_{ij}^2, \\
    \max \mathit{LM} & = & \max_{i = \underline{i}, \dots, \overline{\imath}} ~
      \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1}
      \sum_{j = 1, \dots, k} {\bm B}(\hat {\bm \theta})_{ij}^2.
\end{eqnarray*}

\medskip

\textbf{Critical values:} Analytically for $\mathit{DM}$. Otherwise by
direct simulation or further refined simulation techniques.

\end{frame}

\begin{frame}
\frametitle{Score-based tests: Ordinal variables}

\textbf{Test statistics:} Aggregation along ordinal variables $v$ with $m$ levels.
\begin{eqnarray*}
      \mathit{WDM}_o     & = & \max_{i \in \{i_1, \dots, i_{m-1} \}} ~ \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1/2}      
                             \max_{j = 1, \dots, k} | {\bm B}(\hat {\bm \theta})_{ij} |,\\
      \max \mathit{LM}_o & = & \max_{i \in \{i_1, \dots, i_{m-1} \}} ~
                             \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1}
                             \sum_{j = 1, \dots, k} {\bm B}(\hat {\bm \theta})_{ij}^2,
\end{eqnarray*}
where $i_1, \dots, i_{m-1}$ are the numbers of observations in each category.

\medskip

\textbf{Critical values:} For $\mathit{WDM}_o$ directly from a multivariate normal distribution.
For $\max \mathit{LM}_o$ via simulation.

\end{frame}

\begin{frame}
\frametitle{Score-based tests: Categorical variables}

\textbf{Test statistic:} Aggregation within the $m$ (unordered) categories of $v$.
\[
  \mathit{LM}_{uo} = \sum_{\ell = 1, \dots, m} \sum_{j = 1, \dots, k}
    \left( {\bm B}(\hat {\bm \theta})_{i_\ell j} - {\bm B}(\hat {\bm \theta})_{i_{\ell - 1}j} \right)^2,
\]

\medskip

\textbf{Critical values:} From a $\chi^2$ distribution (as usual).

\medskip

\textbf{Asymptotically equivalent:} LR test.

\end{frame}

\subsection{Illustration}

\begin{frame}
\frametitle{Illustration}

\textbf{Software:} In R system for statistical computing.
\begin{itemize}
  \item \emph{strucchange} implements this general framwork for
    parameter instability tests.
  \item Object-oriented implementation that can be applied to many
    model classes, including \emph{lavaan} objects for CFA models.
\end{itemize}

\medskip

\textbf{Data:}
\begin{itemize}
  \item Application of adult gratitude scale to $n=1401$ youth aged 10--19 years.
  \item GQ-6 scale has five Likert scale items with seven points each.
  \item Assess the factor loadings of a one-factor model.
  \item Question: Measurement invariance across six age groups?
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration}

\textbf{Packages:}

<<packages, results=hide>>=
library("lavaan")
library("strucchange")
@

\medskip

\textbf{Data:} Omitting incomplete cases.

<<data>>=
data("YouthGratitude", package = "psychotools")
compcases <- apply(YouthGratitude[, 4:28], 1,
  function(x) all(x %in% 1:9))
yg <- YouthGratitude[compcases, ]
@

\medskip

\textbf{Estimation:} One-factor CFA with loadings restricted to be equal
across age groups.

<<models>>=
gq6cfa <- cfa("f1 =~ gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
  data = yg, group = "agegroup", meanstructure = TRUE,
  group.equal = "loadings")
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Illustration}

\textbf{Measurement invariance tests:}

<<wdmo, strip.white=all, fig=TRUE, height=4.5, width=5, include=FALSE>>=
sctest(gq6cfa, order.by = yg$agegroup, parm = 1:4,
  vcov = "info", functional = "WDMo", plot = TRUE)
@

<<maxlmo, strip.white=all, fig=TRUE, height=4.5, width=5, include=FALSE>>=
sctest(gq6cfa, order.by = yg$agegroup, parm = 1:4,
  vcov = "info", functional = "maxLMo", plot = TRUE)
@

Both tests reflect only moderate parameter instability across age groups
and do not show significant violations of measurement invariance at 5\% level.

\end{frame}

\begin{frame}
\frametitle{Illustration}

\vspace*{-0.6cm}

\centering
\includegraphics[width=0.8\textwidth]{slides-wdmo}

\end{frame}

\begin{frame}
\frametitle{Illustration}

\vspace*{-0.6cm}

\centering
\includegraphics[width=0.8\textwidth]{slides-maxlmo}

\end{frame}

\subsection{Summary}

\begin{frame}
\frametitle{Summary}

\begin{itemize}
  \item General score-based test framework for assessing measurement
    invariance in parametric psychometric models.
  \item Assessment is along some variable $v$ which can be continuous,
    ordinal, or categorical.
  \item Tests can be seen as generalizations of the Lagrange multiplier
    (or score) test.
  \item Computation of critical values might require simulation from
    certain stochastic processes (Brownian bridges).
  \item Easy-to-use implementation available in R package \emph{strucchange}.
  \item Can be re-used in model-based recursive partitioning.
\end{itemize}

\end{frame}

\subsection{References}

\begin{frame}
\frametitle{References}

\small

Merkle EC, Zeileis A (2013).
  \dquote{Tests of Measurement Invariance without Subgroups: A Generalization of Classical Methods.}
  \emph{Psychometrika}, \textbf{78}(1), 59--82.
  \doi{10.1007/s11336-012-9302-4}

\medskip

Merkle EC, Fan J, Zeileis A (2014).
  \dquote{Testing for Measurement Invariance with Respect to an Ordinal Variable.}
  \emph{Psychometrika}, \textbf{79}(4), 569--584.
  \doi{10.1007/s11336-013-9376-7}

\medskip

Wang T, Merkle EC, Zeileis A (2014).
  \dquote{Score-Based Tests of Measurement Invariance: Use in Practice.}
  \emph{Frontiers in Psychology}, \textbf{5}(438).
  \doi{10.3389/fpsyg.2014.00438}

%% \medskip
%% 
%% Strobl C, Julia Kopf, Zeileis A (2015).
%%   \dquote{Rasch Trees: A New Method for Detecting Differential Item Functioning in the Rasch Model.}
%%   \textit{Psychometrika}. \textbf{80}(2), 289--316.
%%   \doi{10.1007/s11336-013-9388-3}

\end{frame}

\end{document}