### 2015-12-04 ucl2015/slides.Rnw
presentation materials for seminar at UCL
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@388 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[11pt,compress,t]{beamer}
\usetheme{Z}
\usepackage{amsfonts,amstext,amsmath,bm}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}

%% need no \usepackage{Sweave}
\definecolor{InputColor}{rgb}{0,0,0.3}
\definecolor{OutputColor}{rgb}{0.2,0.2,0.2}

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("lavaan")
library("strucchange")
library("psychotree")
library("psychotools")
set.seed(1090)
@

\begin{document}

\title{{\fontsize{14.7}{16}\selectfont Score-Based Tests of Measurement Invariance with Respect to Continuous and Ordinal Variables}}
\author{Achim Zeileis, Edgar C. Merkle, Ting Wang}
\URL{http://eeecon.uibk.ac.at/~zeileis/}
\lecture{Score-Based Tests of Measurement Invariance with Respect to Continuous and Ordinal Variables}

\subsection{Overview}

\begin{frame}
\frametitle{Overview}

\begin{itemize}
  \item Motivation
  \item Framework
  \item Score-based tests
  \begin{itemize}
    \item Continuous variables
    \item Ordinal variables
    \item Categorial variables
  \end{itemize}
  \item Software
  \item Illustrations
\end{itemize}

\end{frame}

\subsection{Motivation}

\begin{frame}
\frametitle{Motivation}

\textbf{Psychometric models:} Typically measure latent scales based on certain
manifest variables, e.g., item response theory (IRT) models or confirmatory
factor analysis (CFA).

\medskip

\textbf{Crucial assumption:} Measurement invariance (MI). Otherwise observed differences in
scales cannot be reliably attributed to the latent variable that the model
purports to measure.

\medskip

\textbf{Parameter stability:} In parametric models, the MI assumption
corresponds to stability of parameters across all possible subgroups.

\medskip

\textbf{Inference:} The typical approach for assessing MI is
\begin{itemize}
  \item to split the data into reference and focal groups,
  \item assess the stability of selected parameters (all or only a subset) across these groups
  \item by means of standard tests: likelihood ratio (LR), Wald, or Lagrange multiplier
    (LM or score) tests.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Motivation}

\textbf{Problems:}
\begin{itemize}
  \item Subgroups have to be formed in advance.
  \item Continuous variables are often categorized into groups in an ad hoc way
    (e.g., splitting at the median).
  \item In ordinal variables the ordering of the categories is often not exploited
    -- assessing only if at least one group differs from the others.
  \item When likelihood ratio or Wald tests are employed, the model has to be
    fitted to each subgroup which can become numerically challenging and computationally
    intensive.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Motivation}

\textbf{Idea:}
\begin{itemize}
  \item Generalize the LM test.
  \item Thus, the model only has to be fitted once under the MI assumption to the full data set.
  \item Catpure model deviations along a variable that is suspected to cause MI violations.
  \item Exploit ordering to assess if there is (at least) one split so that the
    model parameters before and after the split differ.
  \item The split does \emph{not} have to be known or guessed in advance.
\end{itemize}

\medskip

\textbf{Illustration:} CFA for artificial data.
\begin{itemize}
  \item Model with two latent scales (verbal and math).
  \item Three manifest variables for each scale.
  \item Violation of MI for the math loadings along the age of the subjects.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Motivation: CFA for age $\le 16$}

\vspace*{-0.6cm}

\includegraphics[width=0.95\textwidth]{../psychoco2012/model_under16.pdf}

\end{frame}

\begin{frame}
\frametitle{Motivation: CFA for age $> 16$}

\vspace*{-0.6cm}

\includegraphics[width=0.95\textwidth]{../psychoco2012/model_over16.pdf}

\end{frame}

\subsection{Framework}

\begin{frame}
\frametitle{Framework}

\textbf{Model:} Based on log-likelihood $\ell(\cdot)$ for
$p$-dimensional observations $\bm{x}_i$ ($i = 1, \dots, n$)
and $k$-dimensional parameter ${\bm \theta}$.

\medskip

\textbf{Estimation:} Maximum likelihood.
\[
  \hat{\bm{\theta}} ~=~ \argmax_{\bm{\theta}} \sum_{i = 1}^n \ell(\bm{\theta}; \bm{x}_i).
\]

\medskip

\textbf{Equivalently:} Solve first order conditions
\[
  \sum_{i=1}^{n} {\bm s}(\hat{\bm{\theta}}; \bm{x}_i) ~=~ 0,    
\]
where the score function is the partial derivative of the casewise
likelihood contributions w.r.t.\ the parameters $\bm{\theta}$.
\[
  {\bm s}({\bm \theta}; \bm{x}_i) ~=~ \left(
    \frac{\partial \ell({\bm \theta}; \bm{x}_i)}{\partial \theta_1},
    \dots,
    \frac{\partial \ell({\bm \theta}; \bm{x}_i)}{\partial \theta_k}
  \right)^\top.
\]

\end{frame}

\begin{frame}
\frametitle{Framework}

\textbf{Assumption:} Distribution/likelihood of $\bm{x}_i$ depends
only on the latent scales (through the parameters $\bm{\theta}$) --
but not on any other variable $v_i$.

\medskip

\textbf{Alternative view:} Parameters $\bm{\theta}$ do not depend
any such variable $v_i$. Hence assess for $i = 1, \ldots, n$
\begin{eqnarray*}
    H_0: {\bm \theta}_i & = & {\bm \theta}_0, \\
    H_1: {\bm \theta}_i & = & {\bm \theta}(v_i).
\end{eqnarray*}

\medskip

\textbf{Special case:} Two subgroups resulting from one split point $\nu$.
\begin{eqnarray*}
  H_1^*: {\bm \theta}_i & = & \left\{ \begin{array}{ll}
    {\bm \theta}^{(A)} & \mbox{if } v_i \le \nu \\
    {\bm \theta}^{(B)} & \mbox{if } v_i >   \nu
  \end{array} \right.
\end{eqnarray*}

\medskip

\textbf{Tests:} LR/Wald/LM tests can be easily employed if pattern $\bm{\theta}(v_i)$
is known, specifically for $H_1^*$ with fixed split point $\nu$.

\end{frame}

\begin{frame}
\frametitle{Framework}

\textbf{For unknown split points:} Compute LR/Wald/LM tests for each possible
split point $v_1 \le v_2 \le \dots \le v_n$ and reject if the maximum statistic is large.

\medskip

\textbf{Caution:} By maximally selecting the test statistic different critical
values are required (not from a $\chi^2$ distribution)!

\medskip

\textbf{Illustration:} Assess all $k^* = 19$ model parameters from the artificial
CFA example along the continuous variable age ($v_i$).

\end{frame}

\begin{frame}
\frametitle{Framework}

\vspace*{-1.2cm}

\includegraphics[width=\textwidth]{../psychoco2012/paper-lm-lr-19.pdf}

\end{frame}

\begin{frame}
\frametitle{Framework}

\textbf{Note:} For the maxLM test the parameters $\hat{\bm \theta}$ only have
to be estimated once. Only the model scores ${\bm s}(\hat{\bm \theta}; \bm{x}_i)$
have to be aggregated differently for each split point.

\medskip

\textbf{More generally:} Consider a class of tests that
assesses whether the model ``deviations'' ${\bm s}(\hat{\bm \theta}; \bm{x}_i)$
depend on $v_i$. This can consider only a subset $k^*$ of all $k$
parameters/scores or try to capture other patterns than $H_1^*$.

\end{frame}

\begin{frame}
\frametitle{Score-based tests}

\textbf{Fluctuation process:} Capture fluctuations in the cumulative
sum of the scores ordered by the variable $v$.
\[
  {\bm B}(t; \hat {\bm \theta}) ~=~ \hat {\bm I}^{-1/2} n^{-1/2}
    \sum_{i = 1}^{\lfloor n \cdot t \rfloor} {\bm s}(\hat {\bm \theta}; \bm{x}_{(i)})
  \qquad (0 \le t \le 1).
\]
\begin{itemize}
  \item $\hat{\bm I}$ -- estimate of the information matrix.
  \item $t$ -- proportion of data ordered by $v$.
  \item $\lfloor n \cdot t \rfloor$ -- integer part of $n \cdot t$.
  \item $x_{(i)}$ -- observation with the $i$-th smallest value of the
    variable $v$.
\end{itemize}

\medskip

\textbf{Functional central limit theorem:} Under $H_0$ convergence to a (continuous) Brownian bridge process
${\bm B}(\cdot; \hat {\bm \theta}) ~\overset{d}{\rightarrow}~ {\bm B}^{0}(\cdot)$,
from which critical values can be obtained -- either analytically or by simulation.

\end{frame}

\begin{frame}
\frametitle{Score-based tests: Continuous variables}

\textbf{Test statistics:} The empirical process can be viewed as a matrix
${\bm B}(\hat {\bm \theta})_{ij}$ with rows $i = 1, \dots, n$ (observations)
and columns $j = 1, \dots, k$ (parameters). This can be aggregated to scalar
test statistics along continuous the variable $v$.
\begin{eqnarray*}
    \mathit{DM}      & = & \max_{i = 1,\dots, n} \max_{j = 1, \dots, k} | {\bm B}(\hat {\bm \theta})_{ij} | \\
    \mathit{CvM}     & = & n^{-1} \sum_{i = 1,\dots, n} \sum_{j = 1, \dots, k} {\bm B}(\hat {\bm \theta})_{ij}^2, \\
    \max \mathit{LM} & = & \max_{i = \underline{i}, \dots, \overline{\imath}} ~
      \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1}
      \sum_{j = 1, \dots, k} {\bm B}(\hat {\bm \theta})_{ij}^2.
\end{eqnarray*}

\medskip

\textbf{Critical values:} Analytically for $\mathit{DM}$. Otherwise by
direct simulation or further refined simulation techniques.

\end{frame}

\begin{frame}
\frametitle{Score-based tests: Continuous variables}

<<gefp, echo=FALSE, results=hide>>=
load("mz.rda")
source("../www/mz.R")
gefp_3_info <- gefp(mz_fit, fit = NULL, order.by = mz_data$age, parm = 1:3,
  vcov = info.mzfit, sandwich = FALSE)
gefp_19_info <- gefp(mz_fit, fit = NULL, order.by = mz_data$age,
  vcov = info.mzfit, sandwich = FALSE)
@

\setkeys{Gin}{width=1.05\textwidth}
\hspace*{-0.4cm}%
<<gefp-dmax, echo=FALSE, results=hide, fig=TRUE, height=4.5, width=9>>=
par(mfcol = c(1, 2))
plot(gefp_3_info,  functional = maxBB, main = "DM, k* = 3", xlab = "Age")
plot(gefp_19_info, functional = maxBB, main = "DM, k* = 19", xlab = "Age")
@

%plot(gefp_3_info,  functional = meanL2BB, main = "CvM, k* = 3", xlab = "Age")
%plot(gefp_3_info,  functional = supLM(0.1), main = "max LM, k* = 3", xlab = "Age")
%plot(gefp_19_info, functional = meanL2BB, main = "CvM, k* = 19", xlab = "Age")
%plot(gefp_19_info, functional = supLM(0.1), main = "max LM, k* = 19", xlab = "Age")

\end{frame}

\begin{frame}
\frametitle{Score-based tests: Continuous variables}

\setkeys{Gin}{width=1.05\textwidth}
\hspace*{-0.4cm}%
<<gefp-cvm, echo=FALSE, results=hide, fig=TRUE, height=4.5, width=9>>=
par(mfcol = c(1, 2))
plot(gefp_3_info,  functional = meanL2BB, main = "CvM, k* = 3", xlab = "Age")
plot(gefp_19_info, functional = meanL2BB, main = "CvM, k* = 19", xlab = "Age")
@

%plot(gefp_3_info,  functional = supLM(0.1), main = "max LM, k* = 3", xlab = "Age")
%plot(gefp_19_info, functional = supLM(0.1), main = "max LM, k* = 19", xlab = "Age")

\end{frame}

\begin{frame}
\frametitle{Score-based tests: Continuous variables}

\setkeys{Gin}{width=1.05\textwidth}
\hspace*{-0.4cm}%
<<gefp-maxlm, echo=FALSE, results=hide, fig=TRUE, height=4.5, width=9>>=
par(mfcol = c(1, 2))
plot(gefp_3_info,  functional = supLM(0.1), main = "max LM, k* = 3", xlab = "Age")
plot(gefp_19_info, functional = supLM(0.1), main = "max LM, k* = 19", xlab = "Age")
@

\end{frame}

\begin{frame}
\frametitle{Score-based tests: Ordinal variables}

\textbf{Test statistics:} Aggregation along ordinal variables $v$ with $m$ levels.
\begin{eqnarray*}
      \mathit{WDM}_o     & = & \max_{i \in \{i_1, \dots, i_{m-1} \}} ~ \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1/2}      
                             \max_{j = 1, \dots, k} | {\bm B}(\hat {\bm \theta})_{ij} |,\\
      \max \mathit{LM}_o & = & \max_{i \in \{i_1, \dots, i_{m-1} \}} ~
                             \left\{ \frac{i}{n} \left( 1 - \frac{i}{n} \right) \right\}^{-1}
                             \sum_{j = 1, \dots, k} {\bm B}(\hat {\bm \theta})_{ij}^2,
\end{eqnarray*}
where $i_1, \dots, i_{m-1}$ are the numbers of observations in each category.

\medskip

\textbf{Critical values:} For $\mathit{WDM}_o$ directly from a multivariate normal distribution.
For $\max \mathit{LM}_o$ via simulation.

\end{frame}

\begin{frame}
\frametitle{Score-based tests: Categorical variables}

\textbf{Test statistic:} Aggregation within the $m$ (unordered) categories of $v$.
\[
  \mathit{LM}_{uo} = \sum_{\ell = 1, \dots, m} \sum_{j = 1, \dots, k}
    \left( {\bm B}(\hat {\bm \theta})_{i_\ell j} - {\bm B}(\hat {\bm \theta})_{i_{\ell - 1}j} \right)^2,
\]

\medskip

\textbf{Critical values:} From a $\chi^2$ distribution (as usual).

\medskip

\textbf{Asymptotically equivalent:} LR test.

\end{frame}

\subsection{Software}

\begin{frame}
\frametitle{Software}

\textbf{R packages:}
\begin{itemize}
  \item \emph{strucchange} implements this general framwork for
    parameter instability tests.
  \item Object-oriented implementation that can be applied to many
    model classes, including \emph{lavaan} objects for CFA models.
  \item Other psychometric models that cooperate with \emph{strucchange}
    are provided in \emph{psychotools}, e.g., IRT models (Rasch, partial
    credit, rating scale), Bradley-Terry models for paired comparisons,
    and multinomial processing tree models.
  \item Model-based recursive partitioning based on the general
    parameter instability tests are provided in \emph{partykit}.
  \item Adaptation to psychometric models in \emph{psychotree}.
\end{itemize}

\end{frame}

\subsection{CFA: Youth gratitude}

\begin{frame}
\frametitle{CFA: Youth gratitude}

\textbf{Question:} Does measurement invariance hold across age groups
when an adult gratitude scale is applied to youth subjects?

\medskip

\textbf{Source:} Froh \emph{et al.} (2011, \emph{Psychological Assessment}).
  ``Measuring Gratitude in Youth: Assessing the Psychometric Properties of
  Adult Gratitude Scales in Children and Adolescents.''

\medskip

\textbf{Data:}
\begin{itemize}
  \item GQ-6 gratitude scale with five Likert scale items of seven points each.
  \item Application to $n=1401$ youth aged 10--19 years (six age groups).
  \item Assess the factor loadings of a one-factor model.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{CFA: Youth gratitude}

\textbf{Packages:}

<<packages, results=hide>>=
library("lavaan")
library("strucchange")
@

\medskip

\textbf{Data:} Omitting incomplete cases.

<<data>>=
data("YouthGratitude", package = "psychotools")
compcases <- apply(YouthGratitude[, 4:28], 1,
  function(x) all(x %in% 1:9))
yg <- YouthGratitude[compcases, ]
@

\medskip

\textbf{Estimation:} One-factor CFA with loadings restricted to be equal
across age groups.

<<models>>=
gq6_cfa <- cfa("f1 =~ gq6_1 + gq6_2 + gq6_3 + gq6_4 + gq6_5",
  data = yg, group = "agegroup", meanstructure = TRUE,
  group.equal = "loadings")
@

\end{frame}

\begin{frame}[fragile]
\frametitle{CFA: Youth gratitude}

\textbf{Measurement invariance tests:}

<<wdmo, strip.white=all, fig=TRUE, height=4.5, width=5, include=FALSE>>=
sctest(gq6_cfa, order.by = yg$agegroup, parm = 1:4,
  vcov = "info", functional = "WDMo", plot = TRUE)
@

<<maxlmo, strip.white=all, fig=TRUE, height=4.5, width=5, include=FALSE>>=
sctest(gq6_cfa, order.by = yg$agegroup, parm = 1:4,
  vcov = "info", functional = "maxLMo", plot = TRUE)
@

Both tests reflect only moderate parameter instability across age groups
and do not show significant violations of measurement invariance at 5\% level.

\end{frame}

\begin{frame}
\frametitle{CFA: Youth gratitude}

\vspace*{-0.6cm}

\centering
\includegraphics[width=0.8\textwidth]{slides-wdmo}

\end{frame}

\begin{frame}
\frametitle{CFA: Youth gratitude}

\vspace*{-0.6cm}

\centering
\includegraphics[width=0.8\textwidth]{slides-maxlmo}

\end{frame}

\subsection{IRT: Examining exams}

\begin{frame}
\frametitle{IRT: Examining exams}

\textbf{Question:} Does measurement invariance hold in a Rasch model for single-choice exam results?

\medskip

\textbf{Source:} Mathematics for first-year business and economics students at Universit\"at Innsbruck.
Online tests (conducted in OpenOLAT) and written exams for 500--1,000 students per semester.

\medskip

\textbf{Data:} Individual results from an end-term exam.
\begin{itemize}
  \item 729 students (out of 941 registered).
  \item 13~single-choice items with five answer alternatives,
    covering the basics of analysis, linear algebra, financial mathematics.
  \item Two groups with partially different item pools (on the same topics). Individual versions of
    items generated via \emph{exams}.
  \item Correctly solved items yield 100\% of associated points. Items without correct solution can
    either be unanswered (0\%) or with an incorrect answer ($-25\%$). Only considered as binary here.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{IRT: Examining exams}

\textbf{Packages:}

<<packages, results=hide>>=
library("psychotools")
library("psychotree")
@

\medskip

\textbf{Data:} Load, select first group, and exclude extreme scorers.

<<data>>=
load("MathExam.rda")
mex <- subset(MathExam, group == 1 & nsolved > 0 & nsolved < 13)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{IRT: Examining exams}

<<solved-plot, eval=FALSE>>=
plot(mex$solved)
@

\vspace*{-0.6cm}

\setkeys{Gin}{width=1.05\textwidth}
\hspace*{-0.4cm}%
<<solved-plot-fig, echo=FALSE, results=hide, fig=TRUE, height=6.2, width=10>>=
<<solved-plot>>
@

\end{frame}

\begin{frame}[fragile]
\frametitle{IRT: Examining exams}

<<credits-plot, eval=FALSE>>=
plot(mex$credits)
@

\vspace*{-0.6cm}

\setkeys{Gin}{width=1.05\textwidth}
\hspace*{-0.4cm}%
<<credits-plot-fig, echo=FALSE, results=hide, fig=TRUE, height=6.2, width=10>>=
<<credits-plot>>
@

\end{frame}

\begin{frame}[fragile]
\frametitle{IRT: Examining exams}

<<raschmodel-plot, eval=FALSE>>=
mex_rasch <- raschmodel(mex$solved)
plot(mex_rasch, type = "profile")
@

\vspace*{-0.6cm}

\setkeys{Gin}{width=1.05\textwidth}
\hspace*{-0.4cm}%
<<raschmodel-plot-fig, echo=FALSE, results=hide, fig=TRUE, height=6.2, width=10>>=
<<raschmodel-plot>>
@

\end{frame}

\begin{frame}[fragile]
\frametitle{IRT: Examining exams}

<<raschmodel-piplot, eval=FALSE>>=
plot(mex_rasch, type = "piplot")
@

\vspace*{-0.4cm}

\setkeys{Gin}{width=1.05\textwidth}
\hspace*{-0.4cm}%
<<raschmodel-piplot-fig, echo=FALSE, results=hide, fig=TRUE, height=6.2, width=10>>=
<<raschmodel-piplot>>
@

\end{frame}

\begin{frame}[fragile]
\frametitle{IRT: Examining exams}

\textbf{Measurement invariance tests:}

<<raschmodel-maxlm, strip.white=all, fig=TRUE, height=6.5, width=7.5, include=FALSE>>=
sctest(mex_rasch, order.by = jitter(mex$tests),
  vcov = "info", functional = "maxLM", plot = TRUE)
@

<<raschmodel-maxlmo, strip.white=all, fig=TRUE, height=6.5, width=7.5, include=FALSE>>=
mex$otests <- cut(mex$tests, breaks = c(0, 14:24, 26),
  ordered = TRUE, labels = c("<= 14", 15:24, ">= 25"))
sctest(mex_rasch, order.by = mex$otests,
  vcov = "info", functional = "maxLMo", plot = TRUE)
@

Clear violation of measurement invariance: Students that performed poorly in the
previous online tests have a different item profile.

\end{frame}

\begin{frame}
\frametitle{IRT: Examining exams}

\vspace*{-0.6cm}

\centering
\includegraphics[width=0.8\textwidth]{slides-raschmodel-maxlm}

\end{frame}

\begin{frame}
\frametitle{IRT: Examining exams}

\vspace*{-0.6cm}

\centering
\includegraphics[width=0.8\textwidth]{slides-raschmodel-maxlmo}

\end{frame}

\begin{frame}[fragile]
\frametitle{IRT: Examining exams}

<<raschtree>>=
mex_tree <- raschtree(solved ~ otests + attempt + semester + study,
  data = mex, vcov = "info", ordinal = "L2")
@

\vspace*{-0.3cm}

\setkeys{Gin}{width=1.05\textwidth}
\hspace*{-0.4cm}%
<<raschtree-fig, echo=FALSE, results=hide, fig=TRUE, height=6.2, width=10>>=
plot(mex_tree)
@

\end{frame}

\subsection{Paired comparisons: Modeling topmodels}

\begin{frame}[fragile]
\frametitle{Paired comparisons: Modeling topmodels}

\textbf{Question:} Does measurement invariance hold for a Bradley-Terry
preference scaling of attractiveness?

\bigskip

\textbf{Source:} Strobl, Wickelmaier, Zeileis
(2010, \textit{Journal of Educational and Behavioral Statistics}).
``Accounting for Individual Differences in Bradley-Terry Models by Means of Recursive Partitioning.''

\bigskip

\textbf{Data:}
\begin{itemize}
  \item Paired comparisons of attractiveness for
        \emph{Germany's Next Topmodel~2007} finalists:
	Barbara, Anni, Hana, Fiona, Mandy, Anja.
  \item Survey with 192~respondents at Universit{\"a}t T{\"u}bingen.
  \item Available covariates: Gender, age, familiarty with the TV~show.
  \item Familiarity assessed by yes/no questions:
        (1)~Do you recognize the women?/Do you know the show?
	(2)~Did you watch it regularly?
	(3)~Did you watch the final show?/Do you know who won?
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Paired comparisons: Modeling topmodels}

\vspace*{-0.7cm}
\begin{center}
\includegraphics[width=0.25\textwidth]{Barbara.jpg} \hspace*{0.1cm}
\includegraphics[width=0.25\textwidth]{Anni.jpg} \hspace*{0.1cm}
\includegraphics[width=0.25\textwidth]{Hana.jpg}

\vspace*{0.1cm}

\includegraphics[width=0.25\textwidth]{Fiona.jpg} \hspace*{0.1cm}
\includegraphics[width=0.25\textwidth]{Mandy.jpg} \hspace*{0.1cm}
\includegraphics[width=0.25\textwidth]{Anja.jpg}
\end{center}

\end{frame}

\begin{frame}[fragile]
\frametitle{Paired comparisons: Modeling topmodels}

<<Topmodel-MOB>>=
data("Topmodel2007", package = "psychotree")
tm_mob <- bttree(preference ~ ., data = Topmodel2007,
  minsplit = 5, ref = "Barbara")
@

\vspace*{-0.3cm}

\setkeys{Gin}{width=1.05\textwidth}
\hspace*{-0.4cm}%
<<Topmodel-plot, echo=FALSE, fig=TRUE, height=6.2, width=10>>=
plot(tm_mob, abbreviate = 1, yscale = c(0, 0.5))
@

\end{frame}

\begin{frame}
\frametitle{Paired comparisons: Modeling topmodels}

\textbf{Recursively partitioned preferences:} Standardized ranking from Bradley-Terry model.

\bigskip

\begin{tabular}{|r|rrrrrr|} \hline
 & Barbara & Anni & Hana & Fiona & Mandy & Anja \\ \hline
<<Topmodel-table, echo=FALSE, results=tex>>=
tm_tab <- round(worth(tm_mob), digits = 2)
tm_tab <- cbind(rownames(tm_tab), format(tm_tab, nsmall = 2))
cat(paste(apply(tm_tab, 1, paste, collapse = " & "), collapse = "\\\\\n"))
@
\\ \hline
\end{tabular}

\end{frame}

\subsection{Summary}

\begin{frame}
\frametitle{Summary}

\begin{itemize}
  \item General score-based test framework for assessing measurement
    invariance in parametric psychometric models.
  \item Assessment is along some variable $v$ which can be continuous,
    ordinal, or categorical.
  \item Tests can be seen as generalizations of the Lagrange multiplier
    test.
  \item Computation of critical values might require simulation from
    certain stochastic processes (Brownian bridges).
  \item Easy-to-use implementation available in R package \emph{strucchange}.
  \item Can be re-used in model-based recursive partitioning in R packages
    \emph{partykit} and \emph{psychotree}.
\end{itemize}

\bigskip

\textbf{Acknowledgments:} This work was supported by National Science Foundation grant SES-1061334.

\end{frame}

\subsection{References}

\begin{frame}
\frametitle{References}

\footnotesize

Merkle EC, Zeileis A (2013).
  \dquote{Tests of Measurement Invariance without Subgroups: A Generalization of Classical Methods.}
  \emph{Psychometrika}, \textbf{78}(1), 59--82.
  \doi{10.1007/s11336-012-9302-4}

\medskip

Merkle EC, Fan J, Zeileis A (2014).
  \dquote{Testing for Measurement Invariance with Respect to an Ordinal Variable.}
  \emph{Psychometrika}, \textbf{79}(4), 569--584.
  \doi{10.1007/s11336-013-9376-7}

\medskip

Wang T, Merkle EC, Zeileis A (2014).
  \dquote{Score-Based Tests of Measurement Invariance: Use in Practice.}
  \emph{Frontiers in Psychology}, \textbf{5}(438).
  \doi{10.3389/fpsyg.2014.00438}

\bigskip

Strobl C, Julia Kopf, Zeileis A (2015).
  \dquote{Rasch Trees: A New Method for Detecting Differential Item Functioning in the Rasch Model.}
  \textit{Psychometrika}. \textbf{80}(2), 289--316.
  \doi{10.1007/s11336-013-9388-3}

\medskip

Strobl C, Wickelmaier F, Zeileis A (2011).
  \dquote{Accounting for Individual Differences in Bradley-Terry Models by Means of Recursive Partitioning.}
  \textit{Journal of Educational and Behavioral Statistics}, \textbf{36}(2), 135--153.
  \doi{10.3102/1076998609359791}

\medskip

Zeileis A, Hothorn T, Hornik K (2008).
 \dquote{Model-Based Recursive Partitioning.}
  \textit{Journal of Computational and Graphical Statistics},
  \textbf{17}(2), 492--514.
  \doi{10.1198/106186008X319331}

\end{frame}

\end{document}